---
title: | 
 | Quantification of SNP Allelic Differences (Bermuda vs Domestic): 
 | 2-Sided Wilcoxon T-tests
author: "Alec Steep"
date: "02/14/2022"
always_allow_html: true
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
    highlight: zenburn
    css: ../style/style.css
---

# CONFIGURATIONS
```{r setup, message=FALSE, results='hide', warning = FALSE, echo = TRUE}
# Set the working directory and tool paths on your local computer.
WD <- "/Users/Alec/Documents/bermuda/20220214_vcf-annotation_steep" 
# Set the gsutil path

knitr::opts_knit$set(root.dir=WD)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = FALSE)
```

##### CSS Top Styling
```{r CSS Top Styling}
write_css = FALSE
if(write_css){
  writeLines("td, th { padding : 3px } th { background-color:white; color:black; border:1px solid black; text-align:center } td {color:black; border:1px solid black; word-wrap:break-word; white-space:nowrap; overflow: hidden; text-overflow: ellipsis; max-width:300px; text-align:left}", con=file.path(normalizePath(WD), "style/style.css"))
}
```

# Goals of Analysis
- Quantification of SNP allelic Ddfferences (Bermuda vs Domestic)
- Compare SNP position and VEP annotation with allelic differences
- Discover General Themes (Genes/Variant Function/Mutational Frequency) within selective sweep regions

# Takeaways
- SNP data support predicted selective sweep regions
- Some selective sweep regions demonstrate simultaneous
  - differences in allele frequencies between (Bermuda and Domestic), and
  - strong correlation between position and allele frequency
- Nearly all potential “driver” mutations within selective sweep regions reside in non-coding regulatory regions

# Prepare Environment & Load Data

## Setup the Environment
```{r Setup the Environment, message=FALSE, results='hide', warning = FALSE, echo = TRUE}
################################################################################
##### Resources and Dependencies ###############################################
################################################################################
# Whether to knit document and display data
knit_time = TRUE
save_bool = FALSE

#BiocManager::install("ensemblVEP")

# Load dependencies
pacs...man <- c("tidyverse","kableExtra","devtools","glue","lubridate","gt",
                "rtracklayer","glue","impute","multtest","liftOver")
for(pac in pacs...man){
  suppressWarnings(suppressPackageStartupMessages(library(pac, character.only = TRUE)))
}

#browseVignettes("ensemblVEP")
############################################################
##### Functions ############################################
############################################################

# Name functions
select <- dplyr::select
map <- purrr::map
desc <- dplyr::desc
arrange <- dplyr::arrange
melt <- reshape2::melt
mutate <- dplyr::mutate
glue <- glue::glue

# Global options
options(dplyr.print_max = 100)
options(scipen=10000)

# Colors
redblue100<-rgb(read.table(paste0(WD,'/colors/redblue100.txt'),sep='\t',row.names=1,header=T))
```

## Source functions
```{r}
source(glue("{WD}/functions/20220214_LoadVepVcf.R"))
```

### Source of Germline Variant Calls
Germline variants were received from Maria Luisa Martin Cerezo (Marisa) and Dominic Wright on February 14, 2022.

### Source of Annotation Files
- GalGal4 Reference Genome File
  - https://hgdownload.soe.ucsc.edu/goldenPath/galGal4/bigZips/
- GalGal5 Reference Genome File and dictionary
  - https://hgdownload.soe.ucsc.edu/goldenPath/galGal5/bigZips/
- Gg4 to Gg5 chain file
  - https://hgdownload.soe.ucsc.edu/goldenPath/galGal4/liftOver/

# Docker Specific Analyses (performed outside of R)

###  Perform a LiftOver of the g.vcf files via picard

#### Pull the image if neccessary
docker pull broadinstitute/gatk:latest

docker needs permission to bind to directory on mac, see link: https://docs.docker.com/desktop/mac/

#### start up the GATK container with the working directory mounted in
docker run -v /Users/Alec/Documents/bermuda/20220214_vcf-annotation_steep/data:/gatk/bermuda_data -it broadinstitute/gatk:latest
gatk CreateSequenceDictionary -R ./bermuda_data/galGal5.fa -O ./bermuda_data/galGal5.dict
https://www.biostars.org/p/46331/

#### Index the file to extract certain regions
gatk IndexFeatureFile \
--input ./bermuda_data/bermuda_old_kauai_publicseq_filtered_snps.vcf

#### Select variants over selective sweep regions
gatk SelectVariants \
--variant ./bermuda_data/bermuda_old_kauai_publicseq_filtered_snps.vcf \
--output ./bermuda_data/bermuda_old_kauai_publicseq_filtered_snps_sweeps.vcf \
--intervals chr1:150,340,000-150,420,000 \
--intervals chr2:94,120,000-94,220,000 \
--intervals chr2:123,740,000-123,800,000 \
--intervals chr2:142,940,000-143,220,000 \
--intervals chr3:52,840,000-52,900,000 \
--intervals chr3:92,900,000-92,960,000 \
--intervals chr5:3,700,000-3,960,000 \
--intervals chr7:19,320,000-19,380,000

### Liftover the GG4 VCF calls to GG5

picard script was adjusted to allow for 40g RAM if neccessary
Command takes 1 hr to run
gatk LiftoverVcf \
--java-options "-Xmx40G" \
-I ./bermuda_data/bermuda_old_kauai_publicseq_filtered_snps_sweeps.vcf \
-O ./bermuda_data/20220214_bermuda-kauai-public-filtered-gg5-snps-liftover-sweeps_steep.vcf \
--CHAIN ./bermuda_data/galGal4ToGalGal5.over.chain \
--REJECT ./bermuda_data/20220214_bermuda-kauai-public-filtered-gg5-snps-liftover-rejects-sweeps_steep.vcf \
-R ./bermuda_data/galGal5.fa

### Seperate samples via GATK (docker image)
The Genome Analysis Toolkit (GATK) v4.2.5.0
HTSJDK Version: 2.24.1
Picard Version: 2.25.4

samples=("DRS042875" "DRS042876" "DRS042877" "DRS042878" "DRS042879" "P4806_126" "P4806_128" "P4806_131" "P4806_132" "P4806_133" "P4806_134" "P4806_135" "P4806_136" "P4806_137" "P4806_139" "P4806_140" "P4806_141" "P4806_142" "P4806_143" "P4806_144" "P4806_157" "P4806_158" "P4806_159" "P4806_160" "P4806_161" "P4806_162" "SRP022583" "SRS1014597" "SRS1014598" "SRS1014599" "SRS1014600" "SRS1014601" "SRS1014602" "SRS1014603" "SRS1014604" "SRS1014605" "SRS1275748" "SRS1275749" "SRS1275750" "SRS1275751" "SRS1275752" "SRS1275753" "SRS1275754" "SRS1275755" "SRS426963" "SRS524482" "SRS524483" "SRS524484" "SRS524485" "SRS524486" "SRS524487" "SRS524488" "SRS524489" "SRS589235" "SRS589236" "SRS589237" "SRS589238" "SRS589239" "SRS589240" "SRS589241" "SRS589242" "SRS589243" "SRS589244" "SRS589245" "SRS589246" "SRS589247" "SRS589248" "SRS589249" "SRS589250" "SRS589251" "SRS589252" "SRS589253" "SRS589254" "SRS589255" "SRS589256" "SRS589257" "SRS589258" "SRS589259" "SRS589260" "SRS589261" "SRS589262" "SRS589263" "SRS589265" "SRS589266" "SRS810965" "SRS811020" "SRS811021" "SRS811022" "SRS811023" "SRS811024" "SRS811025" "SRS811026" "SRS811027" "ugc_586_1" "ugc_586_10" "ugc_586_11" "ugc_586_12" "ugc_586_13" "ugc_586_14" "ugc_586_15" "ugc_586_16" "ugc_586_17" "ugc_586_18" "ugc_586_19" "ugc_586_2" "ugc_586_20" "ugc_586_21" "ugc_586_22" "ugc_586_23" "ugc_586_25" "ugc_586_3" "ugc_586_4" "ugc_586_5" "ugc_586_6" "ugc_586_7" "ugc_586_8" "ugc_586_9")

for sample in ${samples[@]}; do
echo ${sample};
gatk SelectVariants \
--variant ./bermuda_data/20220214_bermuda-kauai-public-filtered-gg5-snps-liftover-sweeps_steep.vcf \
--output ./bermuda_data/20220214_${sample}-gg5-snps-sweeps_steep.vcf \
--sample-name ${sample}
done

#### Exit the docker image
exit

### Annotate SNPs with VEP annotations

#### Pull the docker image for vep for galgal5
docker pull steepale/vep_galgal5:release_92.1
#### Interactively run the image
docker run -v /Users/Alec/Documents/bermuda/20220214_vcf-annotation_steep/data:/opt/vep/src/ensembl-vep/bermuda_data -it steepale/vep_galgal5:release_92.1

#### List the samples to be annotated (in future, just annotate one sample, all sample share all variants in seperated VCF format)
samples=("DRS042875" "DRS042876" "DRS042877" "DRS042878" "DRS042879" "P4806_126" "P4806_128" "P4806_131" "P4806_132" "P4806_133" "P4806_134" "P4806_135" "P4806_136" "P4806_137" "P4806_139" "P4806_140" "P4806_141" "P4806_142" "P4806_143" "P4806_144" "P4806_157" "P4806_158" "P4806_159" "P4806_160" "P4806_161" "P4806_162" "SRP022583" "SRS1014597" "SRS1014598" "SRS1014599" "SRS1014600" "SRS1014601" "SRS1014602" "SRS1014603" "SRS1014604" "SRS1014605" "SRS1275748" "SRS1275749" "SRS1275750" "SRS1275751" "SRS1275752" "SRS1275753" "SRS1275754" "SRS1275755" "SRS426963" "SRS524482" "SRS524483" "SRS524484" "SRS524485" "SRS524486" "SRS524487" "SRS524488" "SRS524489" "SRS589235" "SRS589236" "SRS589237" "SRS589238" "SRS589239" "SRS589240" "SRS589241" "SRS589242" "SRS589243" "SRS589244" "SRS589245" "SRS589246" "SRS589247" "SRS589248" "SRS589249" "SRS589250" "SRS589251" "SRS589252" "SRS589253" "SRS589254" "SRS589255" "SRS589256" "SRS589257" "SRS589258" "SRS589259" "SRS589260" "SRS589261" "SRS589262" "SRS589263" "SRS589265" "SRS589266" "SRS810965" "SRS811020" "SRS811021" "SRS811022" "SRS811023" "SRS811024" "SRS811025" "SRS811026" "SRS811027" "ugc_586_1" "ugc_586_10" "ugc_586_11" "ugc_586_12" "ugc_586_13" "ugc_586_14" "ugc_586_15" "ugc_586_16" "ugc_586_17" "ugc_586_18" "ugc_586_19" "ugc_586_2" "ugc_586_20" "ugc_586_21" "ugc_586_22" "ugc_586_23" "ugc_586_25" "ugc_586_3" "ugc_586_4" "ugc_586_5" "ugc_586_6" "ugc_586_7" "ugc_586_8" "ugc_586_9")

#### Annotate samples
for sample in ${samples[@]}; do
echo ${sample};
vep \
--fork 4 \
-v \
-i ./bermuda_data/20220214_${sample}-gg5-snps-sweeps_steep.vcf \
-o ./bermuda_data/20220214_${sample}-gg5-snps-sweeps-vep_steep.vcf \
--vcf \
--cache \
--species gallus_gallus \
--force_overwrite \
--sift b \
--protein \
--domains \
--ccds \
--uniprot \
--tsl \
--appris \
--hgvs
done

#### Exit the Docker image
exit

# Load the annotated data back into R

## Load in the vcf files
Each file takes about 1 minute to load
```{r}
# Get a list of the samples to be loaded
samples <- c("DRS042875", "DRS042876", "DRS042877", "DRS042878", "DRS042879", "P4806_126", "P4806_128", "P4806_131", "P4806_132", "P4806_133", "P4806_134", "P4806_135", "P4806_136", "P4806_137", "P4806_139", "P4806_140", "P4806_141", "P4806_142", "P4806_143", "P4806_144", "P4806_157", "P4806_158", "P4806_159", "P4806_160", "P4806_161", "P4806_162", "SRP022583", "SRS1014597", "SRS1014598", "SRS1014599", "SRS1014600", "SRS1014601", "SRS1014602", "SRS1014603", "SRS1014604", "SRS1014605", "SRS1275748", "SRS1275749", "SRS1275750", "SRS1275751", "SRS1275752", "SRS1275753", "SRS1275754", "SRS1275755", "SRS426963", "SRS524482", "SRS524483", "SRS524484", "SRS524485", "SRS524486", "SRS524487", "SRS524488", "SRS524489", "SRS589235", "SRS589236", "SRS589237", "SRS589238", "SRS589239", "SRS589240", "SRS589241", "SRS589242", "SRS589243", "SRS589244", "SRS589245", "SRS589246", "SRS589247", "SRS589248", "SRS589249", "SRS589250", "SRS589251", "SRS589252", "SRS589253", "SRS589254", "SRS589255", "SRS589256", "SRS589257", "SRS589258", "SRS589259", "SRS589260", "SRS589261", "SRS589262", "SRS589263", "SRS589265", "SRS589266", "SRS810965", "SRS811020", "SRS811021", "SRS811022", "SRS811023", "SRS811024", "SRS811025", "SRS811026", "SRS811027", "ugc_586_1", "ugc_586_10", "ugc_586_11", "ugc_586_12", "ugc_586_13", "ugc_586_14", "ugc_586_15", "ugc_586_16", "ugc_586_17", "ugc_586_18", "ugc_586_19", "ugc_586_2", "ugc_586_20", "ugc_586_21", "ugc_586_22", "ugc_586_23", "ugc_586_25", "ugc_586_3", "ugc_586_4", "ugc_586_5", "ugc_586_6", "ugc_586_7", "ugc_586_8", "ugc_586_9")

load_bool1 <- FALSE

if(load_bool1){
  i <- 1
  # Load the vep annotated vcf files
  for(i in 1:length(samples)){
    print(samples[i])
    gg5_vcf_vep <- glue("{WD}/data/20220214_{samples[i]}-gg5-snps-sweeps-vep_steep.vcf")
    eval(call("<-", as.name(glue("{samples[i]}_df")), LoadVepVcf(vcf_file = gg5_vcf_vep)))
  }

  # Combine all the data
  all_df1 <- rbind(DRS042875_df, DRS042876_df, DRS042877_df, DRS042878_df, DRS042879_df, P4806_126_df, P4806_128_df, P4806_131_df, P4806_132_df, P4806_133_df, P4806_134_df, P4806_135_df, P4806_136_df, P4806_137_df, P4806_139_df, P4806_140_df, P4806_141_df, P4806_142_df, P4806_143_df, P4806_144_df, P4806_157_df, P4806_158_df, P4806_159_df, P4806_160_df, P4806_161_df, P4806_162_df, SRP022583_df, SRS1014597_df, SRS1014598_df, SRS1014599_df, SRS1014600_df, SRS1014601_df, SRS1014602_df, SRS1014603_df, SRS1014604_df, SRS1014605_df, SRS1275748_df, SRS1275749_df, SRS1275750_df, SRS1275751_df, SRS1275752_df, SRS1275753_df, SRS1275754_df, SRS1275755_df, SRS426963_df, SRS524482_df, SRS524483_df, SRS524484_df, SRS524485_df, SRS524486_df, SRS524487_df, SRS524488_df, SRS524489_df, SRS589235_df, SRS589236_df, SRS589237_df, SRS589238_df, SRS589239_df, SRS589240_df, SRS589241_df, SRS589242_df, SRS589243_df, SRS589244_df, SRS589245_df, SRS589246_df, SRS589247_df, SRS589248_df, SRS589249_df, SRS589250_df, SRS589251_df, SRS589252_df, SRS589253_df, SRS589254_df, SRS589255_df, SRS589256_df, SRS589257_df, SRS589258_df, SRS589259_df, SRS589260_df, SRS589261_df, SRS589262_df, SRS589263_df, SRS589265_df, SRS589266_df, SRS810965_df, SRS811020_df, SRS811021_df, SRS811022_df, SRS811023_df, SRS811024_df, SRS811025_df, SRS811026_df, SRS811027_df, ugc_586_1_df, ugc_586_10_df, ugc_586_11_df, ugc_586_12_df, ugc_586_13_df, ugc_586_14_df, ugc_586_15_df, ugc_586_16_df, ugc_586_17_df, ugc_586_18_df, ugc_586_19_df, ugc_586_2_df, ugc_586_20_df, ugc_586_21_df, ugc_586_22_df, ugc_586_23_df, ugc_586_25_df, ugc_586_3_df, ugc_586_4_df, ugc_586_5_df, ugc_586_6_df, ugc_586_7_df, ugc_586_8_df, ugc_586_9_df)
  
  # Cleanup
  rm(DRS042875_df, DRS042876_df, DRS042877_df, DRS042878_df, DRS042879_df, P4806_126_df, P4806_128_df, P4806_131_df, P4806_132_df, P4806_133_df, P4806_134_df, P4806_135_df, P4806_136_df, P4806_137_df, P4806_139_df, P4806_140_df, P4806_141_df, P4806_142_df, P4806_143_df, P4806_144_df, P4806_157_df, P4806_158_df, P4806_159_df, P4806_160_df, P4806_161_df, P4806_162_df, SRP022583_df, SRS1014597_df, SRS1014598_df, SRS1014599_df, SRS1014600_df, SRS1014601_df, SRS1014602_df, SRS1014603_df, SRS1014604_df, SRS1014605_df, SRS1275748_df, SRS1275749_df, SRS1275750_df, SRS1275751_df, SRS1275752_df, SRS1275753_df, SRS1275754_df, SRS1275755_df, SRS426963_df, SRS524482_df, SRS524483_df, SRS524484_df, SRS524485_df, SRS524486_df, SRS524487_df, SRS524488_df, SRS524489_df, SRS589235_df, SRS589236_df, SRS589237_df, SRS589238_df, SRS589239_df, SRS589240_df, SRS589241_df, SRS589242_df, SRS589243_df, SRS589244_df, SRS589245_df, SRS589246_df, SRS589247_df, SRS589248_df, SRS589249_df, SRS589250_df, SRS589251_df, SRS589252_df, SRS589253_df, SRS589254_df, SRS589255_df, SRS589256_df, SRS589257_df, SRS589258_df, SRS589259_df, SRS589260_df, SRS589261_df, SRS589262_df, SRS589263_df, SRS589265_df, SRS589266_df, SRS810965_df, SRS811020_df, SRS811021_df, SRS811022_df, SRS811023_df, SRS811024_df, SRS811025_df, SRS811026_df, SRS811027_df, ugc_586_1_df, ugc_586_10_df, ugc_586_11_df, ugc_586_12_df, ugc_586_13_df, ugc_586_14_df, ugc_586_15_df, ugc_586_16_df, ugc_586_17_df, ugc_586_18_df, ugc_586_19_df, ugc_586_2_df, ugc_586_20_df, ugc_586_21_df, ugc_586_22_df, ugc_586_23_df, ugc_586_25_df, ugc_586_3_df, ugc_586_4_df, ugc_586_5_df, ugc_586_6_df, ugc_586_7_df, ugc_586_8_df, ugc_586_9_df)
}
```

## Add the sample cohort annotation
Note: Supp annotation file likely has silent error of columns being accidently shuffled
```{r}
# Load the supplementary data file
supp_file <- glue("{WD}/data/20220214_supp-table_steep.txt")
supp_df <- read.table(file = supp_file, sep = '\t', header = TRUE, fill = TRUE)
# Subset file
supp_df <- supp_df %>%
  select(sample_id_2, cohort, dom, abbreviation, pooled, description)

# Filename
all_file1 <- glue("{WD}/data/20220214_all-gg5-snps-sweeps-vep-v1.0_steep.vcf")
save_bool1 <- FALSE
if(save_bool1){
  # Load the all_df1 file
  # Join the supp data to genetic data
  all_df1 <- left_join(x = all_df1, y = supp_df, by = c("sample" = "sample_id_2"))
  # Save the file
  saveRDS(all_df1, all_file1)
}else{
  # The file has already been adjusted and saved; therefore, just load it
  all_df1 <- readRDS(all_file1)
}
```

# Before Filtering SNPs, Examine the distribution of certain quality statistics for each run

## Determine median depth in sweep region per sample
blue line represents a median depth of 3
```{r}
names(all_df1)

# Examine the median depth in a bar plot stratified by pooled and median depth per sample
# One and zero are bollean for pooled or not
all_df1 %>%
  group_by(sample) %>%
  mutate(DP_med = median(DP, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, DP_med, pooled) %>%
  unique() %>%
  mutate(plot_name = ifelse(DP_med >=3, sample, '')) %>%
  ggplot(aes(x = cohort, y = DP_med)) +
  geom_abline(intercept=3, slope = 0, color = 'blue') +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, height = 0, width = 0.1) +
  facet_wrap(~pooled)

all_df1 %>%
  group_by(sample) %>%
  mutate(DP_med = median(DP, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, DP_med, pooled) %>%
  unique() %>%
  mutate(plot_name = ifelse(DP_med >=3, sample, '')) %>%
  ggplot(aes(color = cohort, x = DP_med)) +
  geom_density() +
  facet_wrap(~pooled)

# Determine samples to remove
sam_rm <- all_df1 %>%
  group_by(sample) %>%
  mutate(DP_med = median(DP, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, DP_med, pooled) %>%
  unique() %>%
  filter(DP_med < 3) %>%
  select(sample) %>% unlist() %>% as.character()
sam_rm

```

## Remove samples (if neccssary, else done donwstream)
```{r}
rm_sam_bool1 <- FALSE
if(rm_sam_bool1){
  all_df2 <- all_df1 %>%
    filter(!sample %in% sam_rm)
}else{
  all_df2 <- all_df1
}
```

## DP
Stratified by cohort and whether samples were pooled
```{r}
# Examine cohort depths
all_df2 %>%
  ggplot(aes(x=DP, color = cohort)) + 
  geom_density(adjust =3) +
  xlim(0,50) +
  facet_wrap(~pooled)
```

### Distribution of Hawaii depths
```{r}
all_df2 %>%
  filter(cohort == 'haw') %>%
  ggplot(aes(x=DP, color = sample)) + 
  geom_density(adjust =3) +
  xlim(0,20) +
  facet_wrap(~pooled)
```

### Distribution of Bermuda depths
```{r}
all_df2 %>%
  filter(cohort == 'ber') %>%
  ggplot(aes(x=DP, color = sample)) + 
  geom_density(adjust =3) +
  xlim(0,50) +
  facet_wrap(~pooled)
```

### Distribution of Domestic depths
```{r}
all_df2 %>%
  filter(cohort == 'dom') %>%
  ggplot(aes(x=DP, color = sample)) + 
  geom_density(adjust =3) +
  xlim(0,50) +
  facet_wrap(~pooled) +
  theme(legend.position="none")
```

### Distribution of RJF depths
```{r}
all_df2 %>%
  filter(cohort == 'rjf') %>%
  ggplot(aes(x=DP, color = sample)) + 
  geom_density(adjust =3) +
  xlim(0,50) +
  facet_wrap(~pooled)
```

## GQ
Stratified by cohort and whether samples were pooled
Note: Genotype qualities for hawaii are considerably lower
```{r}
# Examine cohort depths
all_df2 %>%
  ggplot(aes(x=GQ, color = cohort)) + 
  geom_density(adjust =3) +
  xlim(0,50) +
  facet_wrap(~pooled)
```

### Distribution of Hawaii genotype qualities
```{r}
all_df2 %>%
  filter(cohort == 'haw') %>%
  ggplot(aes(x=GQ, color = sample)) + 
  geom_density(adjust =3) +
  xlim(0,20) +
  facet_wrap(~pooled)
```

### Distribution of Bermuda genotype qualities
```{r}
all_df2 %>%
  filter(cohort == 'ber') %>%
  ggplot(aes(x=GQ, color = sample)) + 
  geom_density(adjust =3) +
  xlim(0,50) +
  facet_wrap(~pooled)
```

### Distribution of Domestic genotype qualities
```{r}
all_df2 %>%
  filter(cohort == 'dom') %>%
  ggplot(aes(x=GQ, color = sample)) + 
  geom_density(adjust =3) +
  xlim(0,50) +
  facet_wrap(~pooled) +
  theme(legend.position="none")
```

### Distribution of RJF genotype qualities
```{r}
all_df2 %>%
  filter(cohort == 'rjf') %>%
  ggplot(aes(x=GQ, color = sample)) + 
  geom_density(adjust =3) +
  xlim(0,50) +
  facet_wrap(~pooled)
```

# Add Additional Annotations

## Annotate SNPs with the following (AA, AB, BB, BC)
```{r}
all_df3 <- all_df2 %>%
  group_by(CHROM, POS, REF, ALT) %>%
  mutate(AA = ifelse(GT == './.', NA, 0)) %>%
  mutate(AB = ifelse(GT == './.', NA, 0)) %>%
  mutate(BB = ifelse(GT == './.', NA, 0)) %>%
  mutate(BC = ifelse(GT == './.', NA, 0)) %>%
  mutate(AA = ifelse(GT == '0/0', 1, AA)) %>%
  mutate(AB = ifelse(GT %in% c('0/1','0/2','0/3'), 1, AB)) %>%
  mutate(BB = ifelse(GT %in% c('1/1','2/2','3/3'), 1, BB)) %>%
  mutate(BC = ifelse(GT %in% c('1/2'), 1, BC)) %>%
  ungroup()
```

### Filter legit calls; limited by data from Hawaii
```{r}
# all_df4 <- all_df3 %>%
#   filter(GT != './.') %>%
#   filter(DP >= 3)
```

## Liftover data from GG5 back to GG4 and add annotation
```{r}

# Chain file for liftover
######################
# Install a chain file for liftover
# Chain format: https://genome.ucsc.edu/goldenpath/help/chain.html
# Liftover file downloaded (20220414) from: https://hgdownload.soe.ucsc.edu/goldenPath/galGal5/liftOver/


################################################################################
#####     Perform Liftover      ################################################
################################################################################

# Add a position-specific index
all_df3 <- all_df3 %>%
  arrange(CHROM,POS) %>%
  group_by(CHROM, POS) %>%
  mutate(pos_index = cur_group_id()) %>%
  ungroup() %>%
  arrange(pos_index) %>%
  select(pos_index, names(all_df3))

# Load the data into a GRanges object
names(all_df3)[53] <- "vep_strand"
all_gr3.1 <- makeGRangesFromDataFrame(all_df3,
                         keep.extra.columns = TRUE,
                         ignore.strand=TRUE,
                         seqnames.field='CHROM',
                         start.field = 'POS',
                         end.field = 'POS')

# Add 'chr' to CHROM names
seqlevelsStyle(all_gr3.1) = "UCSC"

# Import the Chain file
GG5to4_chain <- import.chain(glue("{WD}/data/galGal5ToGalGal4.over.chain"))

# Perform liftover (time)
all_gr3.2 <- liftOver(all_gr3.1, GG5to4_chain)
all_gr3.2 <- unlist(all_gr3.2)
# Convert to tibble
all_df3.2 <- as_tibble(all_gr3.2)
# Drop old REF columns
all_df3.2 <- all_df3.2 %>% dplyr::select(-end, -strand, -width)

# Adjust names
names(all_df3.2)[1:2] <- c('CHROM_gg4','POS_gg4')
all_df3.2 <- all_df3.2 %>% select(pos_index, CHROM_gg4, POS_gg4) %>% unique()
# Join information back into all_df3
all_df3 <- left_join(x = all_df3, y = all_df3.2, by = c("pos_index"))
all_df3 <- all_df3  %>%
  select(sample, CHROM, POS, CHROM_gg4, POS_gg4, names(all_df3)[!names(all_df3) %in% c("sample", "CHROM", "POS", "CHROM_gg4", "POS_gg4")])

# Remove memory object
rm(GG5to4_chain)
```

## Convert tidy dataframe to numeric matrix
AA=0, AB=1, BB=2, BC=3
```{r}
all_df5 <- all_df3 %>%
  mutate(CHROM = as.integer(str_remove_all(as.character(CHROM), 'chr' ))) %>%
  mutate(SNP_VAL = case_when(AA == 1 ~ 0,
                             AB == 1 ~ 1,
                             BB == 1 ~ 2,
                             BC == 1 ~ 3)) %>%
  select(sample, CHROM, POS, SNP_VAL,cohort) %>%
  unique()

# Bermuda; chr 1
ber1.1 <- all_df5 %>%
  filter(cohort == 'ber') %>%
  filter(CHROM == 1) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Bermuda; chr 2
ber2.1 <- all_df5 %>%
  filter(cohort == 'ber') %>%
  filter(CHROM == 2) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Bermuda; chr 3
ber3.1 <- all_df5 %>%
  filter(cohort == 'ber') %>%
  filter(CHROM == 3) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Bermuda; chr 5
ber5.1 <- all_df5 %>%
  filter(cohort == 'ber') %>%
  filter(CHROM == 5) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Bermuda; chr 7
ber7.1 <- all_df5 %>%
  filter(cohort == 'ber') %>%
  filter(CHROM == 7) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()

# domestic; chr 1
dom1.1 <- all_df5 %>%
  filter(cohort == 'dom') %>%
  filter(CHROM == 1) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# domestic; chr 2
dom2.1 <- all_df5 %>%
  filter(cohort == 'dom') %>%
  filter(CHROM == 2) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# domestic; chr 3
dom3.1 <- all_df5 %>%
  filter(cohort == 'dom') %>%
  filter(CHROM == 3) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# domestic; chr 5
dom5.1 <- all_df5 %>%
  filter(cohort == 'dom') %>%
  filter(CHROM == 5) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# domestic; chr 7
dom7.1 <- all_df5 %>%
  filter(cohort == 'dom') %>%
  filter(CHROM == 7) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()

# Red Jungelfowl; chr 1
rjf1.1 <- all_df5 %>%
  filter(cohort == 'rjf') %>%
  filter(CHROM == 1) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Red Jungelfowl; chr 2
rjf2.1 <- all_df5 %>%
  filter(cohort == 'rjf') %>%
  filter(CHROM == 2) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Red Jungelfowl; chr 3
rjf3.1 <- all_df5 %>%
  filter(cohort == 'rjf') %>%
  filter(CHROM == 3) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Red Jungelfowl; chr 5
rjf5.1 <- all_df5 %>%
  filter(cohort == 'rjf') %>%
  filter(CHROM == 5) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Red Jungelfowl; chr 7
rjf7.1 <- all_df5 %>%
  filter(cohort == 'rjf') %>%
  filter(CHROM == 7) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()

# Hawaii; chr 1
haw1.1 <- all_df5 %>%
  filter(cohort == 'haw') %>%
  filter(CHROM == 1) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Hawaii; chr 2
haw2.1 <- all_df5 %>%
  filter(cohort == 'haw') %>%
  filter(CHROM == 2) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Hawaii; chr 3
haw3.1 <- all_df5 %>%
  filter(cohort == 'haw') %>%
  filter(CHROM == 3) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Hawaii; chr 5
haw5.1 <- all_df5 %>%
  filter(cohort == 'haw') %>%
  filter(CHROM == 5) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
# Hawaii; chr 7
haw7.1 <- all_df5 %>%
  filter(cohort == 'haw') %>%
  filter(CHROM == 7) %>%
  select(-cohort, -CHROM) %>%
  pivot_wider(names_from = POS, values_from = SNP_VAL) %>%
  column_to_rownames(var = 'sample') %>% as.matrix()
```

### Print Dimensions (round 1)
```{r}
print("Order: ber, dom, rjf, haw")
print('Chromosome 1')
dim(ber1.1)
dim(dom1.1)
dim(rjf1.1)
dim(haw1.1)
print('Chromosome 2')
dim(ber2.1)
dim(dom2.1)
dim(rjf2.1)
dim(haw2.1)
print('Chromosome 3')
dim(ber3.1)
dim(dom3.1)
dim(rjf3.1)
dim(haw3.1)
print('Chromosome 5')
dim(ber5.1)
dim(dom5.1)
dim(rjf5.1)
dim(haw5.1)
print('Chromosome 7')
dim(ber7.1)
dim(dom7.1)
dim(rjf7.1)
dim(haw7.1)

```

### Save the data in an .RData File
```{r}
if(save_bool){
  save(ber1.1, dom1.1, rjf1.1, haw1.1,
     ber2.1, dom2.1, rjf2.1, haw2.1,
     ber3.1, dom3.1, rjf3.1, haw3.1,
     ber5.1, dom5.1, rjf5.1, haw5.1,
     ber7.1, dom7.1, rjf7.1, haw7.1,
     file = glue("{WD}/data/20220214_cohort-matrices.RData"))
}
```

IMPORTANT: This analysis will not consider hawaii birds because they were not used to detect selective sweep regions

# Perform Initial QC

## Negative/Missing Values (by Feature)

### Rule out negative values
```{r}
# Chrome 1
min(ber1.1,na.rm=T)
min(dom1.1,na.rm=T)
min(rjf1.1,na.rm=T)
# Chrome 2
min(ber2.1,na.rm=T)
min(dom2.1,na.rm=T)
min(rjf2.1,na.rm=T)
# Chrome 3
min(ber3.1,na.rm=T)
min(dom3.1,na.rm=T)
min(rjf3.1,na.rm=T)
# Chrome 5
min(ber5.1,na.rm=T)
min(dom5.1,na.rm=T)
min(rjf5.1,na.rm=T)
# Chrome 7
min(ber7.1,na.rm=T)
min(dom7.1,na.rm=T)
min(rjf7.1,na.rm=T)

```

### Examine the Distribution of Missing values in Features (Chrome 1)
```{r}
# ber
ber1.1.f.c0<-apply(ber1.1,2,function(x) sum(is.na(x))) 
plot(ber1.1.f.c0, ylim = c(0,dim(ber1.1)[1]), main = 'ber; chr1'); abline(h = 5, col = 'blue')

# dom
dom1.1.f.c0<-apply(dom1.1,2,function(x) sum(is.na(x))) 
plot(dom1.1.f.c0, ylim = c(0,dim(dom1.1)[1]), main = 'dom; chr1'); abline(h = 10, col = 'blue')

# rjf
rjf1.1.f.c0<-apply(rjf1.1,2,function(x) sum(is.na(x))) 
plot(rjf1.1.f.c0, ylim = c(0,dim(rjf1.1)[1]), main = 'rjf; chr1'); abline(h = 3, col = 'blue')
```

### Remove Features with High Missing Values (Chrom 1)
```{r}
# ber; chr1
rm_n <- sum(ber1.1.f.c0>=5)
ber1.1 <- ber1.1[,ber1.1.f.c0<5]
print(glue("Features removed from ber;chr1: {rm_n}"))
dim(ber1.1)

# dom; chr1
rm_n <- sum(dom1.1.f.c0>=10)
dom1.1 <- dom1.1[,dom1.1.f.c0<10]
print(glue("Features removed from dom;chr1: {rm_n}"))
dim(dom1.1)

# rjf; chr1
rm_n <- sum(rjf1.1.f.c0>=3)
rjf1.1 <- rjf1.1[,rjf1.1.f.c0<3]
print(glue("Features removed from rjf;chr1: {rm_n}"))
dim(rjf1.1)

```

### Examine the Distribution of Missing values in Features (Chrome 2)
```{r}
# ber
ber2.1.f.c0<-apply(ber2.1,2,function(x) sum(is.na(x))) 
plot(ber2.1.f.c0, ylim = c(0,dim(ber2.1)[1]), main = 'ber; chr2'); abline(h = 5, col = 'blue')

# dom
dom2.1.f.c0<-apply(dom2.1,2,function(x) sum(is.na(x))) 
plot(dom2.1.f.c0, ylim = c(0,dim(dom2.1)[1]), main = 'dom; chr2'); abline(h = 10, col = 'blue')

# rjf
rjf2.1.f.c0<-apply(rjf2.1,2,function(x) sum(is.na(x))) 
plot(rjf2.1.f.c0, ylim = c(0,dim(rjf2.1)[1]), main = 'rjf; chr2'); abline(h = 3, col = 'blue')
```

### Remove Features with High Missing Values (Chrom 2)
```{r}
# ber; chr2
rm_n <- sum(ber2.1.f.c0>=5)
ber2.1 <- ber2.1[,ber2.1.f.c0<5]
print(glue("Features removed from ber;chr2: {rm_n}"))
dim(ber2.1)

# dom; chr2
rm_n <- sum(dom2.1.f.c0>=10)
dom2.1 <- dom2.1[,dom2.1.f.c0<10]
print(glue("Features removed from dom;chr2: {rm_n}"))
dim(dom2.1)

# rjf; chr2
rm_n <- sum(rjf2.1.f.c0>=3)
rjf2.1 <- rjf2.1[,rjf2.1.f.c0<3]
print(glue("Features removed from rjf;chr2: {rm_n}"))
dim(rjf2.1)
```

### Examine the Distribution of Missing values in Features (Chrome 3)
```{r}
# ber
ber3.1.f.c0<-apply(ber3.1,2,function(x) sum(is.na(x))) 
plot(ber3.1.f.c0, ylim = c(0,dim(ber3.1)[1]), main = 'ber; chr3'); abline(h = 5, col = 'blue')

# dom
dom3.1.f.c0<-apply(dom3.1,2,function(x) sum(is.na(x))) 
plot(dom3.1.f.c0, ylim = c(0,dim(dom3.1)[1]), main = 'dom; chr3'); abline(h = 10, col = 'blue')

# rjf
rjf3.1.f.c0<-apply(rjf3.1,2,function(x) sum(is.na(x))) 
plot(rjf3.1.f.c0, ylim = c(0,dim(rjf3.1)[1]), main = 'rjf; chr3'); abline(h = 3, col = 'blue')
```

### Remove Features with High Missing Values (Chrom 3)
```{r}
# ber; chr3
rm_n <- sum(ber3.1.f.c0>=5)
ber3.1 <- ber3.1[,ber3.1.f.c0<5]
print(glue("Features removed from ber;chr3: {rm_n}"))
dim(ber3.1)

# dom; chr3
rm_n <- sum(dom3.1.f.c0>=10)
dom3.1 <- dom3.1[,dom3.1.f.c0<10]
print(glue("Features removed from dom;chr3: {rm_n}"))
dim(dom3.1)

# rjf; chr3
rm_n <- sum(rjf3.1.f.c0>=3)
rjf3.1 <- rjf3.1[,rjf3.1.f.c0<3]
print(glue("Features removed from rjf;chr3: {rm_n}"))
dim(rjf3.1)
```

### Examine the Distribution of Missing values in Features (Chrome 5)
```{r}
# ber
ber5.1.f.c0<-apply(ber5.1,2,function(x) sum(is.na(x))) 
plot(ber5.1.f.c0, ylim = c(0,dim(ber5.1)[1]), main = 'ber; chr5'); abline(h = 5, col = 'blue')

# dom
dom5.1.f.c0<-apply(dom5.1,2,function(x) sum(is.na(x))) 
plot(dom5.1.f.c0, ylim = c(0,dim(dom5.1)[1]), main = 'dom; chr5'); abline(h = 10, col = 'blue')

# rjf
rjf5.1.f.c0<-apply(rjf5.1,2,function(x) sum(is.na(x))) 
plot(rjf5.1.f.c0, ylim = c(0,dim(rjf5.1)[1]), main = 'rjf; chr5'); abline(h = 3, col = 'blue')
```

### Remove Features with High Missing Values (Chrom 5)
```{r}
# ber; chr5
rm_n <- sum(ber5.1.f.c0>=5)
ber5.1 <- ber5.1[,ber5.1.f.c0<5]
print(glue("Features removed from ber;chr5: {rm_n}"))
dim(ber5.1)

# dom; chr5
rm_n <- sum(dom5.1.f.c0>=10)
dom5.1 <- dom5.1[,dom5.1.f.c0<10]
print(glue("Features removed from dom;chr5: {rm_n}"))
dim(dom5.1)

# rjf; chr5
rm_n <- sum(rjf5.1.f.c0>=3)
rjf5.1 <- rjf5.1[,rjf5.1.f.c0<3]
print(glue("Features removed from rjf;chr5: {rm_n}"))
dim(rjf5.1)
```

### Examine the Distribution of Missing values in Features (Chrome 7)
```{r}
# ber
ber7.1.f.c0<-apply(ber7.1,2,function(x) sum(is.na(x))) 
plot(ber7.1.f.c0, ylim = c(0,dim(ber7.1)[1]), main = 'ber; chr7'); abline(h = 5, col = 'blue')

# dom
dom7.1.f.c0<-apply(dom7.1,2,function(x) sum(is.na(x))) 
plot(dom7.1.f.c0, ylim = c(0,dim(dom7.1)[1]), main = 'dom; chr7'); abline(h = 10, col = 'blue')

# rjf
rjf7.1.f.c0<-apply(rjf7.1,2,function(x) sum(is.na(x))) 
plot(rjf7.1.f.c0, ylim = c(0,dim(rjf7.1)[1]), main = 'rjf; chr7'); abline(h = 3, col = 'blue')
```

### Remove Features with High Missing Values (Chrom 7)
```{r}
# ber; chr7
rm_n <- sum(ber7.1.f.c0>=5)
ber7.1 <- ber7.1[,ber7.1.f.c0<5]
print(glue("Features removed from ber;chr7: {rm_n}"))
dim(ber7.1)

# dom; chr7
rm_n <- sum(dom7.1.f.c0>=10)
dom7.1 <- dom7.1[,dom7.1.f.c0<10]
print(glue("Features removed from dom;chr7: {rm_n}"))
dim(dom7.1)

# rjf; chr7
rm_n <- sum(rjf7.1.f.c0>=3)
rjf7.1 <- rjf7.1[,rjf7.1.f.c0<3]
print(glue("Features removed from rjf;chr7: {rm_n}"))
dim(rjf7.1)
```

## Missing Values (by Sample)
Note: We want to make sure that we remove the same samples from each Chromosome, but we will examine the missingness stratified across all chromosomes before making the final decision

### Examine the Distribution of Missing values in Samples (Chrome 1)
```{r}
# ber
ber1.1.s.c0<-apply(ber1.1,1,function(x) sum(is.na(x))) 
plot(ber1.1.s.c0, ylim = c(0,dim(ber1.1)[2]), main = 'ber; chr1'); abline(h = dim(ber1.1)[2]/10, col = 'blue')

# dom
dom1.1.s.c0<-apply(dom1.1,1,function(x) sum(is.na(x))) 
plot(dom1.1.s.c0, ylim = c(0,dim(dom1.1)[2]), main = 'dom; chr1'); abline(h = dim(dom1.1)[2]/10, col = 'blue')

# rjf
rjf1.1.s.c0<-apply(rjf1.1,1,function(x) sum(is.na(x))) 
plot(rjf1.1.s.c0, ylim = c(0,dim(rjf1.1)[2]), main = 'rjf; chr1'); abline(h = dim(rjf1.1)[2]/10, col = 'blue')
```

### Remove Samples with High Missing Values (Chrom 1)
```{r}
# ber; chr1
rm_n <- sum(ber1.1.s.c0>=dim(ber1.1)[2]/10)
rm_names <- names(ber1.1.s.c0)[!ber1.1.s.c0<dim(ber1.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
ber1.1 <- ber1.1[ber1.1.s.c0<dim(ber1.1)[2]/10, ]
print(glue("Samples removed from ber_chr1: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(ber1.1)

# dom; chr1
rm_n <- sum(dom1.1.s.c0>=dim(dom1.1)[2]/10)
rm_names <- names(dom1.1.s.c0)[!dom1.1.s.c0<dim(dom1.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
dom1.1 <- dom1.1[dom1.1.s.c0<dim(dom1.1)[2]/10, ]
print(glue("Samples removed from dom_chr1: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(dom1.1)

# rjf; chr1
rm_n <- sum(rjf1.1.s.c0>=dim(rjf1.1)[2]/10)
rm_names <- names(rjf1.1.s.c0)[!rjf1.1.s.c0<dim(rjf1.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
rjf1.1 <- rjf1.1[rjf1.1.s.c0<dim(rjf1.1)[2]/10, ]
print(glue("Samples removed from rjf_chr1: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(rjf1.1)
```

### Examine the Distribution of Missing values in Samples (Chrom 2)
```{r}
# ber
ber2.1.s.c0<-apply(ber2.1,1,function(x) sum(is.na(x))) 
plot(ber2.1.s.c0, ylim = c(0,dim(ber2.1)[2]), main = 'ber; chr2'); abline(h = dim(ber2.1)[2]/10, col = 'blue')

# dom
dom2.1.s.c0<-apply(dom2.1,1,function(x) sum(is.na(x))) 
plot(dom2.1.s.c0, ylim = c(0,dim(dom2.1)[2]), main = 'dom; chr2'); abline(h = dim(dom2.1)[2]/10, col = 'blue')

# rjf
rjf2.1.s.c0<-apply(rjf2.1,1,function(x) sum(is.na(x))) 
plot(rjf2.1.s.c0, ylim = c(0,dim(rjf2.1)[2]), main = 'rjf; chr2'); abline(h = dim(rjf2.1)[2]/10, col = 'blue')
```

### Remove Samples with High Missing Values (Chrom 2)
```{r}
# ber; chr2
rm_n <- sum(ber2.1.s.c0>=dim(ber2.1)[2]/10)
rm_names <- names(ber2.1.s.c0)[!ber2.1.s.c0<dim(ber2.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
ber2.1 <- ber2.1[ber2.1.s.c0<dim(ber2.1)[2]/10, ]
print(glue("Samples removed from ber_chr2: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(ber2.1)

# dom; chr2
rm_n <- sum(dom2.1.s.c0>=dim(dom2.1)[2]/10)
rm_names <- names(dom2.1.s.c0)[!dom2.1.s.c0<dim(dom2.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
dom2.1 <- dom2.1[dom2.1.s.c0<dim(dom2.1)[2]/10, ]
print(glue("Samples removed from dom_chr2: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(dom2.1)

# rjf; chr2
rm_n <- sum(rjf2.1.s.c0>=dim(rjf2.1)[2]/10)
rm_names <- names(rjf2.1.s.c0)[!rjf2.1.s.c0<dim(rjf2.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
rjf2.1 <- rjf2.1[rjf2.1.s.c0<dim(rjf2.1)[2]/10, ]
print(glue("Samples removed from rjf_chr2: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(rjf2.1)
```

### Examine the Distribution of Missing values in Samples (Chrom 3)
```{r}
# ber
ber3.1.s.c0<-apply(ber3.1,1,function(x) sum(is.na(x))) 
plot(ber3.1.s.c0, ylim = c(0,dim(ber3.1)[2]), main = 'ber; chr3'); abline(h = dim(ber3.1)[2]/10, col = 'blue')

# dom
dom3.1.s.c0<-apply(dom3.1,1,function(x) sum(is.na(x))) 
plot(dom3.1.s.c0, ylim = c(0,dim(dom3.1)[2]), main = 'dom; chr3'); abline(h = dim(dom3.1)[2]/10, col = 'blue')

# rjf
rjf3.1.s.c0<-apply(rjf3.1,1,function(x) sum(is.na(x))) 
plot(rjf3.1.s.c0, ylim = c(0,dim(rjf3.1)[2]), main = 'rjf; chr3'); abline(h = dim(rjf3.1)[2]/10, col = 'blue')
```

### Remove Samples with High Missing Values (Chrom 3)
```{r}
# ber; chr3
rm_n <- sum(ber3.1.s.c0>=dim(ber3.1)[2]/10)
rm_names <- names(ber3.1.s.c0)[!ber3.1.s.c0<dim(ber3.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
ber3.1 <- ber3.1[ber3.1.s.c0<dim(ber3.1)[2]/10, ]
print(glue("Samples removed from ber_chr3: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(ber3.1)

# dom; chr3
rm_n <- sum(dom3.1.s.c0>=dim(dom3.1)[2]/10)
rm_names <- names(dom3.1.s.c0)[!dom3.1.s.c0<dim(dom3.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
dom3.1 <- dom3.1[dom3.1.s.c0<dim(dom3.1)[2]/10, ]
print(glue("Samples removed from dom_chr3: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(dom3.1)

# rjf; chr3
rm_n <- sum(rjf3.1.s.c0>=dim(rjf3.1)[2]/10)
rm_names <- names(rjf3.1.s.c0)[!rjf3.1.s.c0<dim(rjf3.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
rjf3.1 <- rjf3.1[rjf3.1.s.c0<dim(rjf3.1)[2]/10, ]
print(glue("Samples removed from rjf_chr3: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(rjf3.1)
```

### Examine the Distribution of Missing values in Samples (Chrom 5)
```{r}
# ber
ber5.1.s.c0<-apply(ber5.1,1,function(x) sum(is.na(x))) 
plot(ber5.1.s.c0, ylim = c(0,dim(ber5.1)[2]), main = 'ber; chr5'); abline(h = dim(ber5.1)[2]/10, col = 'blue')

# dom
dom5.1.s.c0<-apply(dom5.1,1,function(x) sum(is.na(x))) 
plot(dom5.1.s.c0, ylim = c(0,dim(dom5.1)[2]), main = 'dom; chr5'); abline(h = dim(dom5.1)[2]/10, col = 'blue')

# rjf
rjf5.1.s.c0<-apply(rjf5.1,1,function(x) sum(is.na(x))) 
plot(rjf5.1.s.c0, ylim = c(0,dim(rjf5.1)[2]), main = 'rjf; chr5'); abline(h = dim(rjf5.1)[2]/10, col = 'blue')
```

### Remove Samples with High Missing Values (Chrom 5)
```{r}
# ber; chr5
rm_n <- sum(ber5.1.s.c0>=dim(ber5.1)[2]/10)
rm_names <- names(ber5.1.s.c0)[!ber5.1.s.c0<dim(ber5.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
ber5.1 <- ber5.1[ber5.1.s.c0<dim(ber5.1)[2]/10, ]
print(glue("Samples removed from ber_chr5: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(ber5.1)

# dom; chr5
rm_n <- sum(dom5.1.s.c0>=dim(dom5.1)[2]/10)
rm_names <- names(dom5.1.s.c0)[!dom5.1.s.c0<dim(dom5.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
dom5.1 <- dom5.1[dom5.1.s.c0<dim(dom5.1)[2]/10, ]
print(glue("Samples removed from dom_chr5: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(dom5.1)

# rjf; chr5
rm_n <- sum(rjf5.1.s.c0>=dim(rjf5.1)[2]/10)
rm_names <- names(rjf5.1.s.c0)[!rjf5.1.s.c0<dim(rjf5.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
rjf5.1 <- rjf5.1[rjf5.1.s.c0<dim(rjf5.1)[2]/10, ]
print(glue("Samples removed from rjf_chr5: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(rjf5.1)
```

### Examine the Distribution of Missing values in Samples (Chrom 7)
```{r}
# ber
ber7.1.s.c0<-apply(ber7.1,1,function(x) sum(is.na(x))) 
plot(ber7.1.s.c0, ylim = c(0,dim(ber7.1)[2]), main = 'ber; chr7'); abline(h = dim(ber7.1)[2]/10, col = 'blue')

# dom
dom7.1.s.c0<-apply(dom7.1,1,function(x) sum(is.na(x))) 
plot(dom7.1.s.c0, ylim = c(0,dim(dom7.1)[2]), main = 'dom; chr7'); abline(h = dim(dom7.1)[2]/10, col = 'blue')

# rjf
rjf7.1.s.c0<-apply(rjf7.1,1,function(x) sum(is.na(x))) 
plot(rjf7.1.s.c0, ylim = c(0,dim(rjf7.1)[2]), main = 'rjf; chr7'); abline(h = dim(rjf7.1)[2]/10, col = 'blue')
```

### Remove Samples with High Missing Values (Chrom 7)
```{r}
# ber; chr7
rm_n <- sum(ber7.1.s.c0>=dim(ber7.1)[2]/10)
rm_names <- names(ber7.1.s.c0)[!ber7.1.s.c0<dim(ber7.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
ber7.1 <- ber7.1[ber7.1.s.c0<dim(ber7.1)[2]/10, ]
print(glue("Samples removed from ber_chr7: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(ber7.1)

# dom; chr7
rm_n <- sum(dom7.1.s.c0>=dim(dom7.1)[2]/10)
rm_names <- names(dom7.1.s.c0)[!dom7.1.s.c0<dim(dom7.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
dom7.1 <- dom7.1[dom7.1.s.c0<dim(dom7.1)[2]/10, ]
print(glue("Samples removed from dom_chr7: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(dom7.1)

# rjf; chr7
rm_n <- sum(rjf7.1.s.c0>=dim(rjf7.1)[2]/10)
rm_names <- names(rjf7.1.s.c0)[!rjf7.1.s.c0<dim(rjf7.1)[2]/10]
if(length(rm_names) == 0){rm_names <- 'none'}
rjf7.1 <- rjf7.1[rjf7.1.s.c0<dim(rjf7.1)[2]/10, ]
print(glue("Samples removed from rjf_chr7: {rm_n}; {paste0(rm_names, collapse = ',')}"))
dim(rjf7.1)
```

# Subset only the shared samples and shared features between matrices
Note: Finished matrices should have the same dimensions for each chromosome
Note: All samples are the same between chromosomes

## Shared Features

### Shared Features (Chrom 1)
```{r}
# Collect the shared features
shared_p1 <- colnames(ber1.1)[colnames(ber1.1) %in% colnames(dom1.1)]
shared_p2 <- colnames(rjf1.1)[colnames(rjf1.1) %in% colnames(ber1.1)]
shared_p <- shared_p1[shared_p1 %in% shared_p2]

# Create shared samples and features
ber1.1 <- ber1.1[, shared_p]
dom1.1 <- dom1.1[, shared_p]
rjf1.1 <- rjf1.1[, shared_p]
```

### Shared Features (Chrom 2)
```{r}
# Collect the shared features
shared_p1 <- colnames(ber2.1)[colnames(ber2.1) %in% colnames(dom2.1)]
shared_p2 <- colnames(rjf2.1)[colnames(rjf2.1) %in% colnames(ber2.1)]
shared_p <- shared_p1[shared_p1 %in% shared_p2]

# Create shared samples and features
ber2.1 <- ber2.1[, shared_p]
dom2.1 <- dom2.1[, shared_p]
rjf2.1 <- rjf2.1[, shared_p]
```

### Shared Features (Chrom 3)
```{r}
# Collect the shared features
shared_p1 <- colnames(ber3.1)[colnames(ber3.1) %in% colnames(dom3.1)]
shared_p2 <- colnames(rjf3.1)[colnames(rjf3.1) %in% colnames(ber3.1)]
shared_p <- shared_p1[shared_p1 %in% shared_p2]

# Create shared samples and features
ber3.1 <- ber3.1[, shared_p]
dom3.1 <- dom3.1[, shared_p]
rjf3.1 <- rjf3.1[, shared_p]
```

### Shared Features (Chrom 5)
```{r}
# Collect the shared features
shared_p1 <- colnames(ber5.1)[colnames(ber5.1) %in% colnames(dom5.1)]
shared_p2 <- colnames(rjf5.1)[colnames(rjf5.1) %in% colnames(ber5.1)]
shared_p <- shared_p1[shared_p1 %in% shared_p2]

# Create shared samples and features
ber5.1 <- ber5.1[, shared_p]
dom5.1 <- dom5.1[, shared_p]
rjf5.1 <- rjf5.1[, shared_p]
```

### Shared Features (Chrom 7)
```{r}
# Collect the shared features
shared_p1 <- colnames(ber7.1)[colnames(ber7.1) %in% colnames(dom7.1)]
shared_p2 <- colnames(rjf7.1)[colnames(rjf7.1) %in% colnames(ber7.1)]
shared_p <- shared_p1[shared_p1 %in% shared_p2]

# Create shared samples and features
ber7.1 <- ber7.1[, shared_p]
dom7.1 <- dom7.1[, shared_p]
rjf7.1 <- rjf7.1[, shared_p]
```

### Print Dimensions (round 2)
```{r}
print("Order: ber, dom, rjf, haw")
print('Chromosome 1')
dim(ber1.1)
dim(dom1.1)
dim(rjf1.1)
print('Chromosome 2')
dim(ber2.1)
dim(dom2.1)
dim(rjf2.1)
print('Chromosome 3')
dim(ber3.1)
dim(dom3.1)
dim(rjf3.1)
print('Chromosome 5')
dim(ber5.1)
dim(dom5.1)
dim(rjf5.1)
print('Chromosome 7')
dim(ber7.1)
dim(dom7.1)
dim(rjf7.1)

```

# Imputation

## Confirm the New Distribution of Missing Features

### Missing Features (Chrom 1)
```{r}
# ber
ber1.1.f.c0<-apply(ber1.1,2,function(x) sum(is.na(x))) 
plot(ber1.1.f.c0, ylim = c(0,dim(ber1.1)[2]))

# dom
dom1.1.f.c0<-apply(dom1.1,2,function(x) sum(is.na(x))) 
plot(dom1.1.f.c0, ylim = c(0,dim(dom1.1)[2]))

# rjf
rjf1.1.f.c0<-apply(rjf1.1,2,function(x) sum(is.na(x))) 
plot(rjf1.1.f.c0, ylim = c(0,dim(rjf1.1)[2]))
```

### Missing Features (Chrom 2)
```{r}
# ber
ber2.1.f.c0<-apply(ber2.1,2,function(x) sum(is.na(x))) 
plot(ber2.1.f.c0, ylim = c(0,dim(ber2.1)[2]))

# dom
dom2.1.f.c0<-apply(dom2.1,2,function(x) sum(is.na(x))) 
plot(dom2.1.f.c0, ylim = c(0,dim(dom2.1)[2]))

# rjf
rjf2.1.f.c0<-apply(rjf2.1,2,function(x) sum(is.na(x))) 
plot(rjf2.1.f.c0, ylim = c(0,dim(rjf2.1)[2]))
```

### Missing Features (Chrom 3)
```{r}
# ber
ber3.1.f.c0<-apply(ber3.1,2,function(x) sum(is.na(x))) 
plot(ber3.1.f.c0, ylim = c(0,dim(ber3.1)[2]))

# dom
dom3.1.f.c0<-apply(dom3.1,2,function(x) sum(is.na(x))) 
plot(dom3.1.f.c0, ylim = c(0,dim(dom3.1)[2]))

# rjf
rjf3.1.f.c0<-apply(rjf3.1,2,function(x) sum(is.na(x))) 
plot(rjf3.1.f.c0, ylim = c(0,dim(rjf3.1)[2]))
```

### Missing Features (Chrom 5)
```{r}
# ber
ber5.1.f.c0<-apply(ber5.1,2,function(x) sum(is.na(x))) 
plot(ber5.1.f.c0, ylim = c(0,dim(ber5.1)[2]))

# dom
dom5.1.f.c0<-apply(dom5.1,2,function(x) sum(is.na(x))) 
plot(dom5.1.f.c0, ylim = c(0,dim(dom5.1)[2]))

# rjf
rjf5.1.f.c0<-apply(rjf5.1,2,function(x) sum(is.na(x))) 
plot(rjf5.1.f.c0, ylim = c(0,dim(rjf5.1)[2]))
```

### Missing Features (Chrom 7)
```{r}
# ber
ber7.1.f.c0<-apply(ber7.1,2,function(x) sum(is.na(x))) 
plot(ber7.1.f.c0, ylim = c(0,dim(ber7.1)[2]))

# dom
dom7.1.f.c0<-apply(dom7.1,2,function(x) sum(is.na(x))) 
plot(dom7.1.f.c0, ylim = c(0,dim(dom7.1)[2]))

# rjf
rjf7.1.f.c0<-apply(rjf7.1,2,function(x) sum(is.na(x))) 
plot(rjf7.1.f.c0, ylim = c(0,dim(rjf7.1)[2]))
```

## Imputation: A seperate object will be created with imputed data ... imputated variants will not be considered entirely trustworthy

### Impute Missing values (Chrome 1)
```{r}
#ber
glue("ber: features with missing values (rows) before and after imputation")
ber1.1.2<-impute.knn(ber1.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(ber1.1[,ber1.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(ber1.1.2[, ber1.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(ber1.1.2))}")

#dom
glue("dom: features with missing values (rows) before and after imputation")
dom1.1.2<-impute.knn(dom1.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(dom1.1[,dom1.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(dom1.1.2[, dom1.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(dom1.1.2))}")

#rjf
glue("rjf: features with missing values (rows) before and after imputation")
rjf1.1.2<-impute.knn(rjf1.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(rjf1.1[,rjf1.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(rjf1.1.2[, rjf1.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(rjf1.1.2))}")
```

### Impute Missing values (Chrome 2)
```{r}
#ber
glue("ber: features with missing values (rows) before and after imputation")
ber2.1.2<-impute.knn(ber2.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(ber2.1[,ber2.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(ber2.1.2[, ber2.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(ber2.1.2))}")

#dom
glue("dom: features with missing values (rows) before and after imputation")
dom2.1.2<-impute.knn(dom2.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(dom2.1[,dom2.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(dom2.1.2[, dom2.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(dom2.1.2))}")

#rjf
glue("rjf: features with missing values (rows) before and after imputation")
rjf2.1.2<-impute.knn(rjf2.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(rjf2.1[,rjf2.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(rjf2.1.2[, rjf2.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(rjf2.1.2))}")
```

### Impute Missing values (Chrome 3)
```{r}
#ber
glue("ber: features with missing values (rows) before and after imputation")
ber3.1.2<-impute.knn(ber3.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(ber3.1[,ber3.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(ber3.1.2[, ber3.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(ber3.1.2))}")

#dom
glue("dom: features with missing values (rows) before and after imputation")
dom3.1.2<-impute.knn(dom3.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(dom3.1[,dom3.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(dom3.1.2[, dom3.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(dom3.1.2))}")

#rjf
glue("rjf: features with missing values (rows) before and after imputation")
rjf3.1.2<-impute.knn(rjf3.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(rjf3.1[,rjf3.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(rjf3.1.2[, rjf3.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(rjf3.1.2))}")
```

### Impute Missing values (Chrome 5)
```{r}
#ber
glue("ber: features with missing values (rows) before and after imputation")
ber5.1.2<-impute.knn(ber5.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
try(
  image(as.matrix(ber5.1[,ber5.1.f.c0>0]),col=redblue100,axes=F),
  silent=FALSE)
try(image(as.matrix(ber5.1.2[, ber5.1.f.c0>0]),col=redblue100,axes=F),
  silent=FALSE)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(ber5.1.2))}")

#dom
glue("dom: features with missing values (rows) before and after imputation")
dom5.1.2<-impute.knn(dom5.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(dom5.1[,dom5.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(dom5.1.2[, dom5.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(dom5.1.2))}")

#rjf
glue("rjf: features with missing values (rows) before and after imputation")
rjf5.1.2<-impute.knn(rjf5.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(rjf5.1[,rjf5.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(rjf5.1.2[, rjf5.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(rjf5.1.2))}")
```

### Impute Missing values (Chrome 7)
```{r}
#ber
glue("ber: features with missing values (rows) before and after imputation")
ber7.1.2<-impute.knn(ber7.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
try(image(as.matrix(ber7.1[,ber7.1.f.c0>0]),col=redblue100,axes=F))
try(image(as.matrix(ber7.1.2[, ber7.1.f.c0>0]),col=redblue100,axes=F))
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(ber7.1.2))}")

#dom
glue("dom: features with missing values (rows) before and after imputation")
dom7.1.2<-impute.knn(dom7.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(dom7.1[,dom7.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(dom7.1.2[, dom7.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(dom7.1.2))}")

#rjf
glue("rjf: features with missing values (rows) before and after imputation")
rjf7.1.2<-impute.knn(rjf7.1,k=100)$data
#view the features before and after imputation
par(mfrow=c(2,1),bg="black")
image(as.matrix(rjf7.1[,rjf7.1.f.c0>0]),col=redblue100,axes=F)
image(as.matrix(rjf7.1.2[, rjf7.1.f.c0>0]),col=redblue100,axes=F)
par(mfrow=c(1,1) ,bg="white")
glue("Verified no missing values: {sum(is.na(rjf7.1.2))}")
```

# Create Sample Annotation Matrices and Combine Data

### Create Sample Annotation Matrix (Chrom 1)
```{r}
all_df1$DP1_med <- all_df1$DP2_med <- all_df1$DP3_med <- all_df1$DP5_med <- all_df1$DP7_med <- NA
all_df1$GQ1_med <- all_df1$GQ2_med <- all_df1$GQ3_med <- all_df1$GQ5_med <- all_df1$GQ7_med <- NA
sam_ann_mat1.1 <- all_df1 %>%
  filter(sample %in% c(rownames(ber1.1), rownames(dom1.1), rownames(rjf1.1)) )%>%
  filter(CHROM == 'chr1') %>%
  filter(POS %in% colnames(ber1.1)) %>%
  group_by(sample) %>%
  mutate(DP1.1_med = median(DP, na.rm = TRUE)) %>%
  mutate(GQ1.1_med = median(GQ, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, dom, pooled, DP1.1_med, GQ1.1_med) %>% unique() %>%
  mutate(cohort = case_when(cohort == 'dom' ~ 1,
                            cohort == 'ber' ~ 2,
                            cohort == 'rjf' ~ 3)) %>%
  arrange(cohort, pooled, sample) %>%
  column_to_rownames(var = 'sample') %>%
  as.matrix()

sam_ann_mat1.1
plot(sam_ann_mat1.1[,4],sam_ann_mat1.1[,5])
```

### Create Sample Annotation Matrix (Chrom 2)
```{r}
sam_ann_mat2.1 <- all_df1 %>%
  filter(sample %in% c(rownames(ber2.1), rownames(dom2.1), rownames(rjf2.1)) )%>%
  filter(CHROM == 'chr2') %>%
  filter(POS %in% colnames(ber2.1)) %>%
  group_by(sample) %>%
  mutate(DP2.1_med = median(DP, na.rm = TRUE)) %>%
  mutate(GQ2.1_med = median(GQ, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, dom, pooled, DP2.1_med, GQ2.1_med) %>% unique() %>%
  mutate(cohort = case_when(cohort == 'dom' ~ 1,
                            cohort == 'ber' ~ 2,
                            cohort == 'rjf' ~ 3)) %>%
  arrange(cohort, pooled, sample) %>%
  column_to_rownames(var = 'sample') %>%
  as.matrix()

sam_ann_mat2.1
plot(sam_ann_mat2.1[,4],sam_ann_mat2.1[,5])
```

### Create Sample Annotation Matrix (Chrom 3)
```{r}
sam_ann_mat3.1 <- all_df1 %>%
  filter(sample %in% c(rownames(ber3.1), rownames(dom3.1), rownames(rjf3.1)) )%>%
  filter(CHROM == 'chr3') %>%
  filter(POS %in% colnames(ber3.1)) %>%
  group_by(sample) %>%
  mutate(DP3.1_med = median(DP, na.rm = TRUE)) %>%
  mutate(GQ3.1_med = median(GQ, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, dom, pooled, DP3.1_med, GQ3.1_med) %>% unique() %>%
  mutate(cohort = case_when(cohort == 'dom' ~ 1,
                            cohort == 'ber' ~ 2,
                            cohort == 'rjf' ~ 3)) %>%
  arrange(cohort, pooled, sample) %>%
  column_to_rownames(var = 'sample') %>%
  as.matrix()

sam_ann_mat3.1
plot(sam_ann_mat3.1[,4],sam_ann_mat3.1[,5])
```

### Create Sample Annotation Matrix (Chrom 5)
```{r}
sam_ann_mat5.1 <- all_df1 %>%
  filter(sample %in% c(rownames(ber5.1), rownames(dom5.1), rownames(rjf5.1)) )%>%
  filter(CHROM == 'chr5') %>%
  filter(POS %in% colnames(ber5.1)) %>%
  group_by(sample) %>%
  mutate(DP5.1_med = median(DP, na.rm = TRUE)) %>%
  mutate(GQ5.1_med = median(GQ, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, dom, pooled, DP5.1_med, GQ5.1_med) %>% unique() %>%
  mutate(cohort = case_when(cohort == 'dom' ~ 1,
                            cohort == 'ber' ~ 2,
                            cohort == 'rjf' ~ 3)) %>%
  arrange(cohort, pooled, sample) %>%
  column_to_rownames(var = 'sample') %>%
  as.matrix()

sam_ann_mat5.1
plot(sam_ann_mat5.1[,4],sam_ann_mat5.1[,5])
```

### Create Sample Annotation Matrix (Chrom 7)
```{r}
sam_ann_mat7.1 <- all_df1 %>%
  filter(sample %in% c(rownames(ber7.1), rownames(dom7.1), rownames(rjf7.1)) )%>%
  filter(CHROM == 'chr7') %>%
  filter(POS %in% colnames(ber7.1)) %>%
  group_by(sample) %>%
  mutate(DP7.1_med = median(DP, na.rm = TRUE)) %>%
  mutate(GQ7.1_med = median(GQ, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, dom, pooled, DP7.1_med, GQ7.1_med) %>% unique() %>%
  mutate(cohort = case_when(cohort == 'dom' ~ 1,
                            cohort == 'ber' ~ 2,
                            cohort == 'rjf' ~ 3)) %>%
  arrange(cohort, pooled, sample) %>%
  column_to_rownames(var = 'sample') %>%
  as.matrix()

sam_ann_df7.1 <- all_df1 %>%
  filter(sample %in% c(rownames(ber7.1), rownames(dom7.1), rownames(rjf7.1)) ) %>%
  filter(CHROM == 'chr7') %>%
  filter(POS %in% colnames(ber7.1)) %>%
  group_by(sample) %>%
  mutate(DP7.1_med = median(DP, na.rm = TRUE)) %>%
  mutate(GQ7.1_med = median(GQ, na.rm = TRUE)) %>%
  ungroup() %>%
  select(sample, cohort, dom, pooled, DP7.1_med, GQ7.1_med) %>% unique() %>%
  mutate(cohort = case_when(cohort == 'dom' ~ 1,
                            cohort == 'ber' ~ 2,
                            cohort == 'rjf' ~ 3)) %>%
  arrange(cohort, pooled, sample) %>%
  column_to_rownames(var = 'sample')

sam_ann_mat7.1

plot(sam_ann_mat7.1[,4],sam_ann_mat7.1[,5])
sam_ann_df7.1 %>%
  mutate(cohort = factor(cohort)) %>%
  ggplot(aes(x = DP7.1_med,y=GQ7.1_med, color = cohort)) +
  geom_point()


```

## Combine the cohort matrixes into multiple cohorts and single chromosome 

### Combine cohort matrices (Chrom 1)
```{r}
## Imputed matrices
# Check colnames are in same order
all(colnames(ber1.1.2) == colnames(dom1.1.2))
all(colnames(dom1.1.2) == colnames(rjf1.1.2))

# Create matrix (imputed)
all1.1.2 <- rbind(ber1.1.2, dom1.1.2, rjf1.1.2)

# Rearrange matrix sample order
all1.1.2 <- all1.1.2[rownames(sam_ann_mat1.1),]

## Non-imputed matrices
# Check colnames are in same order
all(colnames(ber1.1) == colnames(dom1.1))
all(colnames(dom1.1) == colnames(rjf1.1))

# Create matrix (imputed)
all1.1 <- rbind(ber1.1, dom1.1, rjf1.1)

# Rearrange matrix sample order
all1.1 <- all1.1[rownames(sam_ann_mat1.1),]

### Z-Transformed matrices
# ber
image(ber1.1.2)
image(log2(ber1.1.2))
image(log2(ber1.1.2+1))

ber.f.mean <- apply(log2(ber1.1.2+1), 2, mean)
ber.f.sd <- apply(log2(ber1.1.2+1), 2, sd)
plot(ber.f.mean,ber.f.sd)

ber1.3a <- ber1.1.2
for (i in 1:dim(ber1.1.2)[1]) {
  ber1.3a[,i]<-(log2(ber1.1.2[,i] + 1)- ber.f.mean[i])/ ber.f.sd[i]
}

# dom
image(dom1.1.2)
image(log2(dom1.1.2))
image(log2(dom1.1.2+1))

dom.f.mean <- apply(log2(dom1.1.2+1), 2, mean)
dom.f.sd <- apply(log2(dom1.1.2+1), 2, sd)
plot(dom.f.mean,dom.f.sd)

dom1.3a <- dom1.1.2
for (i in 1:dim(dom1.1.2)[1]) {
  dom1.3a[,i]<-(log2(dom1.1.2[,i] + 1)- dom.f.mean[i])/ dom.f.sd[i]
}

# rjf
image(rjf1.1.2)
image(log2(rjf1.1.2))
image(log2(rjf1.1.2+1))

rjf.f.mean <- apply(log2(rjf1.1.2+1), 2, mean)
rjf.f.sd <- apply(log2(rjf1.1.2+1), 2, sd)
plot(rjf.f.mean,rjf.f.sd)

rjf1.3a <- rjf1.1.2
for (i in 1:dim(rjf1.1.2)[1]) {
  rjf1.3a[,i]<-(log2(rjf1.1.2[,i] + 1)- rjf.f.mean[i])/ rjf.f.sd[i]
}

## Non-imputed matrices
# Check colnames are in same order
all(colnames(ber1.3a) == colnames(dom1.3a))
all(colnames(dom1.3a) == colnames(rjf1.3a))

# Create matrix (imputed)
all1.3a <- rbind(ber1.3a, dom1.3a, rjf1.3a)

# Rearrange matrix sample order
all1.3a <- all1.3a[rownames(sam_ann_mat1.1),]

```

### Combine cohort matrices (Chrom 2)
```{r}
## Imputed matrices
# Check colnames are in same order
all(colnames(ber2.1.2) == colnames(dom2.1.2))
all(colnames(dom2.1.2) == colnames(rjf2.1.2))

# Create matrix (imputed)
all2.1.2 <- rbind(ber2.1.2, dom2.1.2, rjf2.1.2)

# Rearrange matrix sample order
all2.1.2 <- all2.1.2[rownames(sam_ann_mat2.1),]

## Non-imputed matrices
# Check colnames are in same order
all(colnames(ber2.1) == colnames(dom2.1))
all(colnames(dom2.1) == colnames(rjf2.1))

# Create matrix (imputed)
all2.1 <- rbind(ber2.1, dom2.1, rjf2.1)

# Rearrange matrix sample order
all2.1 <- all2.1[rownames(sam_ann_mat2.1),]
```

### Combine cohort matrices (Chrom 3)
```{r}
## Imputed matrices
# Check colnames are in same order
all(colnames(ber3.1.2) == colnames(dom3.1.2))
all(colnames(dom3.1.2) == colnames(rjf3.1.2))

# Create matrix (imputed)
all3.1.2 <- rbind(ber3.1.2, dom3.1.2, rjf3.1.2)

# Rearrange matrix sample order
all3.1.2 <- all3.1.2[rownames(sam_ann_mat3.1),]

## Non-imputed matrices
# Check colnames are in same order
all(colnames(ber3.1) == colnames(dom3.1))
all(colnames(dom3.1) == colnames(rjf3.1))

# Create matrix (imputed)
all3.1 <- rbind(ber3.1, dom3.1, rjf3.1)

# Rearrange matrix sample order
all3.1 <- all3.1[rownames(sam_ann_mat3.1),]
```

### Combine cohort matrices (Chrom 5)
```{r}
## Imputed matrices
# Check colnames are in same order
all(colnames(ber5.1.2) == colnames(dom5.1.2))
all(colnames(dom5.1.2) == colnames(rjf5.1.2))

# Create matrix (imputed)
all5.1.2 <- rbind(ber5.1.2, dom5.1.2, rjf5.1.2)

# Rearrange matrix sample order
all5.1.2 <- all5.1.2[rownames(sam_ann_mat5.1),]

## Non-imputed matrices
# Check colnames are in same order
all(colnames(ber5.1) == colnames(dom5.1))
all(colnames(dom5.1) == colnames(rjf5.1))

# Create matrix (imputed)
all5.1 <- rbind(ber5.1, dom5.1, rjf5.1)

# Rearrange matrix sample order
all5.1 <- all5.1[rownames(sam_ann_mat5.1),]
```

### Combine cohort matrices (Chrom 7)
```{r}
## Imputed matrices
# Check colnames are in same order
all(colnames(ber7.1.2) == colnames(dom7.1.2))
all(colnames(dom7.1.2) == colnames(rjf7.1.2))

# Create matrix (imputed)
all7.1.2 <- rbind(ber7.1.2, dom7.1.2, rjf7.1.2)

# Rearrange matrix sample order
all7.1.2 <- all7.1.2[rownames(sam_ann_mat7.1),]

## Non-imputed matrices
# Check colnames are in same order
all(colnames(ber7.1) == colnames(dom7.1))
all(colnames(dom7.1) == colnames(rjf7.1))

# Create matrix (imputed)
all7.1 <- rbind(ber7.1, dom7.1, rjf7.1)

# Rearrange matrix sample order
all7.1 <- all7.1[rownames(sam_ann_mat7.1),]
```

# Differential Allelic Frequency Tests: 2-sided, non-parametric t-tests (wilcoxon; data not centered or scaled)
Significance Tests applied to each sweep region seperately

## Sweep Region 1 (chr1:150340000-150420000)

### T-test (dom vs ber) (chr1:150340000-150420000)
```{r}
# Perform a differential genotype test between domestic and bermuda
sam_ann_mat1.1.df <- as.data.frame(sam_ann_mat1.1)
dom_sams1.1 <- sam_ann_mat1.1.df %>%
  filter(cohort == 1) %>% row.names()
ber_sams1.1 <- sam_ann_mat1.1.df %>%
  filter(cohort == 2) %>% row.names()

# T-test
dom1.1.2 <- all1.1.2[dom_sams1.1,]
ber1.1.2 <- all1.1.2[ber_sams1.1,]
dim(dom1.1.2)
dim(ber1.1.2)

t.ber.dom  <- rep(0,dim(all1.1.2)[2])
cls <- c(rep(0, dim(dom1.1.2)[1]), rep(1, dim(ber1.1.2)[1]))
t.ber.dom <- mt.teststat(t(all1.1.2[1:75,]), cls, test="wilcoxon")
qqnorm(t.ber.dom); qqline(t.ber.dom)
qqnorm(t.ber.dom, ylim = c(-6.5,6.5), xlim = c(-4,4)); qqline(t.ber.dom)

```

### Variants with Positive T-score (chr1:150,340,000-150,420,000)
```{r}
# Positive t-scores
t.df.p <- data.frame('P' = colnames(all1.1.2), 't_val' = t.ber.dom) %>%
  filter(t_val > 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.p2 <- t.df.p %>% filter(!is.na(t_val))
t.df.p2 <- t.df.p2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all1.1.2[1:75,], col = redblue100)
image(all1.1.2[1:75,t.df.p2$P], col = redblue100)

# Select the chrom specific vars
ch1_var_df <- all_df3 %>%
  filter(CHROM == 'chr1') %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.p2 <- t.df.p2 %>%
  left_join(y = ch1_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.p <- t.df.p2 %>%
  filter(fdr<=0.10) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr1') %>% 
  filter(POS %in% t.sig.p) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Variants with Negative T-score (chr1:150,340,000-150,420,000)
```{r}
# Negative t-scores
t.df.n <- data.frame('P' = colnames(all1.1.2), 't_val' = t.ber.dom) %>%
  filter(t_val < 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.n2 <- t.df.n %>% filter(!is.na(t_val))
t.df.n2 <- t.df.n2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all1.1.2[1:75,], col = redblue100)
image(all1.1.2[1:75,t.df.n2$P], col = redblue100)

# Select the chrom 2 specific vars
ch1_var_df <- all_df3 %>%
  filter(CHROM == 'chr1') %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.n2 <- t.df.n2 %>%
  left_join(y = ch1_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.n <- t.df.n2 %>%
  filter(fdr<=0.10) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr1') %>% filter(POS %in% t.sig.n) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Visualize the t-scores by location (chr1:150,340,000-150,420,000)
```{r}
# Bind dfs
r1.t.df <- rbind(t.df.p2, t.df.n2) %>%
  mutate(t_sign = ifelse(t_val >0, 'pos', 'neg')) %>%
  mutate(t_val_abs = abs(t_val)) %>%
  mutate(del = ifelse(grepl('del', SIFT), 'del', 'tol'))

# Visualize gg4 position by t scores
r1.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(150340000, 150420000) +
  theme_bw()

r1.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(150340000, 150420000) +
  theme_bw() +
  facet_wrap(~ensembl_geneid + var_type)

# Visualize gg4 position by t scores
r1.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  facet_wrap(~var_type)

# Relavent genes (geneid)
r1.t.df %>%
  filter(ensembl_geneid != '') %>%
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0,1) +
  xlim(150340000, 150420000) +
  facet_wrap(~ensembl_geneid + var_type)


# Deleterious variants
r1.t.df %>%
  filter(grepl('del', SIFT))

# Relavent genes (gene symbol)
# try(
#   r1.t.df %>%
#   filter(ensembl_geneid != '') %>%
#   filter(del == 'del') %>%
#   ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign, shape = del)) +
#   geom_point() +
#   xlim(150340000, 150420000) +
#   ylim(0,1) +
#   facet_wrap(~gene_symbol + var_type)
# )
```

### Examine the variantions of interest (chr1:150340000-150420000)
```{r}
r1.roi <- r1.t.df %>%
  #filter(ensembl_geneid != '') %>%
  filter(fdr <= 0.11) %>%
  select(P) %>% unique() %>% unlist() %>% unname()

# Ann
sidebar<-cbind(cls, cls, cls, cls, cls, cls, cls)
sidebar <- sidebar*3

hmo <- heatmap(all1.1.2[c(dom_sams1.1, dom_sams1.1),r1.roi])$colInd

# Unclustered
image(
  cbind(all1.1.2[c(dom_sams1.1, ber_sams1.1),r1.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr1:150340000-150420000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

# Clustered
clst_mat <- all1.1.2[c(dom_sams1.1, ber_sams1.1),r1.roi]
clst_mat <- clst_mat[,hmo]
image(
  cbind(clst_mat, sidebar),
  col=redblue100,axes=F,main=glue("chr1:150340000-150420000, Cohort Order, Clustered"), asp = 1,
  zlim=c(0,3))
```

## Sweep Region 2 (chr2:94120000-94220000)

### T-test (dom vs ber) (chr2:94,120,000-94,220,000)
```{r}
# Perform a differential genotype test between domestic and bermuda
sam_ann_mat2.r2.df <- as.data.frame(sam_ann_mat2.1)
dom_sams2.r2 <- sam_ann_mat2.r2.df %>%
  filter(cohort == 1) %>% row.names()
ber_sams2.r2 <- sam_ann_mat2.r2.df %>%
  filter(cohort == 2) %>% row.names()

# Collect the region specific snps
r.gg5.pos <- all_df3 %>%
  filter(POS %in% colnames(dom2.1.2)) %>%
  select(CHROM,POS,CHROM_gg4,POS_gg4) %>% unique() %>%
  arrange(POS_gg4) %>%
  filter(POS_gg4 >= 94120000) %>% filter(POS_gg4 <= 94220000) %>%
  select(POS) %>% unlist() %>% unname() %>% as.character()

# Collect region specific locations
dom2.r2.2 <- all2.1.2[dom_sams2.r2,r.gg5.pos]
ber2.r2.2 <- all2.1.2[ber_sams2.r2,r.gg5.pos]
all2.r2.2 <- all2.1.2[c(dom_sams2.r2, ber_sams2.r2),r.gg5.pos]
row.names(all2.r2.2)
dim(dom2.r2.2)
dim(ber2.r2.2)
dim(all2.r2.2)

# T-test
t.ber.dom  <- rep(0,dim(all2.r2.2)[2])
cls <- c(rep(0, dim(dom2.r2.2)[1]), rep(1, dim(ber2.r2.2)[1]))
t.ber.dom <- mt.teststat(t(all2.r2.2), cls, test="wilcoxon")
qqnorm(t.ber.dom); qqline(t.ber.dom)
qqnorm(t.ber.dom, ylim = c(-6.5,6.5), xlim = c(-4,4)); qqline(t.ber.dom)
```

### Variants with Positive T-score (chr2:94,120,000-94,220,000)
```{r}
# Positive t-scores
t.df.p <- data.frame('P' = colnames(all2.r2.2), 't_val' = t.ber.dom) %>%
  filter(t_val > 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.p2 <- t.df.p %>% filter(!is.na(t_val))
t.df.p2 <- t.df.p2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all2.r2.2, col = redblue100)
image(all2.r2.2[,t.df.p2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr2') %>%
  filter(POS_gg4 >= 94120000) %>% filter(POS_gg4 <= 94220000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.p2 <- t.df.p2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.p <- t.df.p2 %>%
  filter(fdr<=0.10) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr2') %>% 
  filter(POS %in% t.sig.p) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Variants with Negative T-score (chr2:94,120,000-94,220,000)
```{r}
# Negative t-scores
t.df.n <- data.frame('P' = colnames(all2.r2.2), 't_val' = t.ber.dom) %>%
  filter(t_val < 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.n2 <- t.df.n %>% filter(!is.na(t_val))
t.df.n2 <- t.df.n2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all2.r2.2, col = redblue100)
image(all2.r2.2[,t.df.n2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr2') %>%
  filter(POS_gg4 >= 94120000) %>% filter(POS_gg4 <= 94220000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.n2 <- t.df.n2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.n <- t.df.n2 %>%
  filter(fdr<=0.10) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr2') %>% filter(POS %in% t.sig.n) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Visualize the t-scores by location (chr2:94120000-94220000)
```{r}
# Bind dfs
r2.t.df <- rbind(t.df.p2, t.df.n2) %>%
  mutate(t_sign = ifelse(t_val >0, 'pos', 'neg')) %>%
  mutate(t_val_abs = abs(t_val)) %>%
  mutate(del = ifelse(grepl('del', SIFT), 'del', 'tol'))

# Visualize gg4 position by t scores
r2.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(94120000,94220000) +
  theme_bw()

r2.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(94120000,94220000) +
  theme_bw() +
  facet_wrap(~ ensembl_geneid + var_type)

# Visualize gg4 position by t scores
r2.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  xlim(94120000,94220000) +
  facet_wrap(~var_type)

# Relavent genes (geneid)
r2.t.df %>%
  filter(ensembl_geneid != '') %>%
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0,1) +
  xlim(94120000,94220000) +
  facet_wrap(~gene_symbol + var_type)
r2.t.df %>%
  #filter(ensembl_geneid != '') %>%
  filter(fdr <= 0.2) %>%
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0.8,1) +
  xlim(94120000,94220000) +
  facet_wrap(~gene_symbol + var_type)

# Deleterious variants
r2.t.df %>%
  filter(fdr <= 0.2) %>%
  filter(gene_symbol == 'CCDC102B') %>%
  filter(var_type == 'missense_variant')

# Relavent genes (gene symbol)
r2.t.df %>%
  filter(ensembl_geneid != '') %>%
  filter(del == 'del') %>%
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign, shape = del)) +
  geom_point() +
  xlim(94120000,94220000) +
  ylim(0,1) +
  facet_wrap(~gene_symbol + var_type)
```

### Examine the variantions of interest (chr2:123740000-123800000)
```{r}
r2.roi <- r2.t.df %>%
  filter(ensembl_geneid != '') %>%
  filter(fdr <= 0.05) %>%
  select(P) %>% unique() %>% unlist() %>% unname()

# Ann
sidebar<-cbind(cls)
sidebar <- sidebar*3

hmo <- heatmap(all2.1.2[c(dom_sams2.r2, ber_sams2.r2),r2.roi])$colInd
hmo_dom <- heatmap(all2.1.2[c(dom_sams2.r2),r2.roi])$rowInd
hmo_ber <- heatmap(all2.1.2[c(ber_sams2.r2),r2.roi])$rowInd

# Unclustered
image(
  cbind(all2.1.2[c(dom_sams2.r2, ber_sams2.r2),r2.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr2:123740000-123800000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

# Clustered
clst_mat <- all2.1.2[c(hmo_dom, hmo_ber),r2.roi]
clst_mat <- clst_mat[,hmo]
image(
  cbind(clst_mat, sidebar),
  col=redblue100,axes=F,main=glue("chr2:123740000-123800000, Cohort Order, Clustered"), asp = 1,
  zlim=c(0,3))
```

## Sweep Region 3 (chr2:123740000-123800000)

### T-test (dom vs ber) (chr2:123740000-123800000)
```{r}
# Perform a differential genotype test between domestic and bermuda
sam_ann_mat2.r2.df <- as.data.frame(sam_ann_mat2.1)
dom_sams2.r2 <- sam_ann_mat2.r2.df %>%
  filter(cohort == 1) %>% row.names()
ber_sams2.r2 <- sam_ann_mat2.r2.df %>%
  filter(cohort == 2) %>% row.names()
dim(dom2.1.2)
# Collect the region specific snps
r.gg5.pos <- all_df3 %>%
  filter(POS %in% colnames(dom2.1.2)) %>%
  select(CHROM,POS,CHROM_gg4,POS_gg4) %>% 
  unique() %>%
  arrange(POS_gg4) %>%
  filter(POS_gg4 >= 123740000) %>% filter(POS_gg4 <= 123800000) %>%
  select(POS) %>% unlist() %>% unname() %>% as.character()

# Collect region specific locations
dom2.r2.2 <- all2.1.2[dom_sams2.r2,r.gg5.pos]
ber2.r2.2 <- all2.1.2[ber_sams2.r2,r.gg5.pos]
all2.r2.2 <- all2.1.2[c(dom_sams2.r2, ber_sams2.r2),r.gg5.pos]
row.names(all2.r2.2)
dim(dom2.r2.2)
dim(ber2.r2.2)
dim(all2.r2.2)

# T-test
t.ber.dom  <- rep(0,dim(all2.r2.2)[2])
cls <- c(rep(0, dim(dom2.r2.2)[1]), rep(1, dim(ber2.r2.2)[1]))
t.ber.dom <- mt.teststat(t(all2.r2.2), cls, test="wilcoxon")
qqnorm(t.ber.dom); qqline(t.ber.dom)
qqnorm(t.ber.dom, ylim = c(-6.5,6.5), xlim = c(-4,4)); qqline(t.ber.dom)
```

### Variants with Positive T-score (chr2:123740000-123800000)
```{r}
# Positive t-scores
t.df.p <- data.frame('P' = colnames(all2.r2.2), 't_val' = t.ber.dom) %>%
  filter(t_val > 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.p2 <- t.df.p %>% filter(!is.na(t_val))
t.df.p2 <- t.df.p2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all2.r2.2, col = redblue100)
image(all2.r2.2[,t.df.p2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr2') %>%
  filter(POS_gg4 >= 123740000) %>% filter(POS_gg4 <= 123800000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.p2 <- t.df.p2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.p <- t.df.p2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr2') %>% 
  filter(POS %in% t.sig.p) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Variants with Negative T-score (chr2:123740000-123800000)
```{r}
# Negative t-scores
t.df.n <- data.frame('P' = colnames(all2.r2.2), 't_val' = t.ber.dom) %>%
  filter(t_val < 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.n2 <- t.df.n %>% filter(!is.na(t_val))
t.df.n2 <- t.df.n2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all2.r2.2, col = redblue100)
image(all2.r2.2[,t.df.n2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr2') %>%
  filter(POS_gg4 >= 123740000) %>% filter(POS_gg4 <= 123800000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.n2 <- t.df.n2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.n <- t.df.n2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr2') %>% filter(POS %in% t.sig.n) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Visualize the t-scores by location (chr2:123740000-123800000)
```{r}
# Bind dfs
r3.t.df <- rbind(t.df.p2, t.df.n2) %>%
  mutate(t_sign = ifelse(t_val >0, 'pos', 'neg')) %>%
  mutate(t_val_abs = abs(t_val)) %>%
  mutate(del = ifelse(grepl('del', SIFT), 'del', 'tol'))

r3.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(123740000,123800000) +
  theme_bw()

r3.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(123740000,123800000) +
  theme_bw() +
  facet_wrap(~ ensembl_geneid + var_type)

# Visualize gg4 position by t scores
r3.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  xlim(123740000,123800000) +
  facet_wrap(~var_type)

# Relavent genes (geneid)
# r3.t.df %>%
#   filter(ensembl_geneid != '') %>%
#   ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
#   geom_point() +
#   ylim(0,1) +
#   xlim(123740000,123800000) +
#   facet_wrap(~ensembl_geneid + var_type)
r3.t.df %>%
  #filter(ensembl_geneid != '') %>%
  filter(fdr <= 0.1) %>%
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0.9,1) +
  xlim(123740000,123800000)
  #facet_wrap(~ensembl_geneid + var_type)
```

### Examine the variantions of interest (chr2:123740000-123800000)
```{r}
r3.roi <- r3.t.df %>%
  #filter(ensembl_geneid != '') %>%
  filter(fdr <= 0.1) %>%
  select(P) %>% unique() %>% unlist() %>% unname()

# Ann
sidebar<-cbind(cls)
sidebar <- sidebar*3

hmo <- heatmap(all2.1.2[c(dom_sams2.r2, ber_sams2.r2),r3.roi])$colInd

# Unclustered
image(
  cbind(all2.1.2[c(dom_sams2.r2, ber_sams2.r2),r3.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr2:123740000-123800000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

clst_mat <- all2.1.2[c(dom_sams2.r2, ber_sams2.r2),r3.roi]
clst_mat <- clst_mat[,hmo]

# Clustered
image(
  cbind(clst_mat, sidebar),
  col=redblue100,axes=F,main=glue("chr2:123740000-123800000, Cohort Order, Clustered"), asp = 1,
  zlim=c(0,3))
```

## Sweep Region 4 (chr2:123740000-123800000)

### T-test (dom vs ber) (chr2:123740000-123800000)
```{r}
# Perform a differential genotype test between domestic and bermuda
sam_ann_mat2.r4.df <- as.data.frame(sam_ann_mat2.1)
dom_sams2.r4 <- sam_ann_mat2.r4.df %>%
  filter(cohort == 1) %>% row.names()
ber_sams2.r4 <- sam_ann_mat2.r4.df %>%
  filter(cohort == 2) %>% row.names()
rjf_sams2.r4 <- sam_ann_mat2.r4.df %>%
  filter(cohort == 3) %>% row.names()

# Collect the region specific snps
r.gg5.pos <- all_df3 %>%
  filter(POS %in% colnames(dom2.1.2)) %>%
  select(CHROM,POS,CHROM_gg4,POS_gg4) %>% unique() %>%
  arrange(POS_gg4) %>%
  filter(POS_gg4 >= 142940000) %>% filter(POS_gg4 <= 143220000) %>%
  select(POS) %>% unlist() %>% unname() %>% as.character()

# Collect region specific locations
dom2.r4.2 <- all2.1.2[dom_sams2.r4,r.gg5.pos]
ber2.r4.2 <- all2.1.2[ber_sams2.r4,r.gg5.pos]
all2.r4.2 <- all2.1.2[c(dom_sams2.r4, ber_sams2.r4),r.gg5.pos]
row.names(all2.r4.2)
dim(dom2.r4.2)
dim(ber2.r4.2)
dim(all2.r4.2)

# T-test
t.ber.dom  <- rep(0,dim(all2.r4.2)[2])
cls <- c(rep(0, dim(dom2.r4.2)[1]), rep(1, dim(ber2.r4.2)[1]))
t.ber.dom <- mt.teststat(t(all2.r4.2), cls, test="wilcoxon")
qqnorm(t.ber.dom); qqline(t.ber.dom)
qqnorm(t.ber.dom, ylim = c(-6.5,6.5), xlim = c(-4,4)); qqline(t.ber.dom)

# Create a sequence of numbers between -10 and 10 incrementing by 0.1.
x <- seq(-4, 4, by = .001)

# Choose the mean as 2.5 and standard deviation as 0.5.
y <- dnorm(x, mean = 0, sd = 1)

plot(x,y)

```

### Variants with Positive T-score (chr2:142940000-143220000)
```{r}
# Positive t-scores
t.df.p <- data.frame('P' = colnames(all2.r4.2), 't_val' = t.ber.dom) %>%
  filter(t_val > 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.p2 <- t.df.p %>% filter(!is.na(t_val))
t.df.p2 <- t.df.p2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))

image(all2.r4.2, col = redblue100)
image(all2.r4.2[,t.df.p2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr2') %>%
  filter(POS_gg4 >= 142940000) %>% filter(POS_gg4 <= 143220000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.p2 <- t.df.p2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.p <- t.df.p2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr2') %>% 
  filter(POS %in% t.sig.p) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Variants with Negative T-score (chr2:142940000-143220000)
```{r}
# Negative t-scores
t.df.n <- data.frame('P' = colnames(all2.r4.2), 't_val' = t.ber.dom) %>%
  filter(t_val < 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.n2 <- t.df.n %>% filter(!is.na(t_val))
t.df.n2 <- t.df.n2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all2.r4.2, col = redblue100)
image(all2.r4.2[,t.df.n2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr2') %>%
  filter(POS_gg4 >= 142940000) %>% filter(POS_gg4 <= 143220000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.n2 <- t.df.n2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.n <- t.df.n2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr2') %>% filter(POS %in% t.sig.n) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Visualize the t-scores by location (chr2:142940000-143220000)
```{r}
# Bind dfs
r4.t.df <- rbind(t.df.p2, t.df.n2) %>%
  mutate(t_sign = ifelse(t_val >0, 'pos', 'neg')) %>%
  mutate(t_val_abs = abs(t_val)) %>%
  mutate(del = ifelse(grepl('del', SIFT), 'del', 'tol'))

r4.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(142940000,143220000) +
  theme_bw()

r4.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(142940000,143220000) +
  theme_bw() +
  facet_wrap(~ ensembl_geneid + var_type)

# Visualize gg4 position by t scores
r4.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  xlim(142940000,143220000) +
  facet_wrap(~var_type)

# Relavent genes (geneid)
r4.t.df %>%
  #filter(ensembl_geneid != '') %>%
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0.95,1) +
  xlim(142940000,143220000) +
  facet_wrap(~var_type)
```

### Examine the variantions of interest (chr2:142940000-143220000)
```{r}
r4.roi <- r4.t.df %>%
  #filter(ensembl_geneid != '') %>%
  filter(fdr <= 0.05) %>%
  select(P) %>% unique() %>% unlist() %>% unname()

# Ann
sidebar<-cbind(cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,
               cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,
               cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls,cls)
sidebar <- sidebar*3

hmo <- heatmap(all2.1.2[c(dom_sams2.r4, ber_sams2.r4),r4.roi])$colInd

# Unclustered
image(
  cbind(all2.1.2[c(dom_sams2.r4, ber_sams2.r4),r4.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr2:142940000-143220000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

# Clustered
clst_mat <- all2.1.2[c(dom_sams2.r4, ber_sams2.r4),r4.roi]
clst_mat <- clst_mat[,hmo]
image(
  cbind(clst_mat, sidebar),
  col=redblue100,axes=F,main=glue("chr2:142940000-143220000, Cohort Order, Clustered"), asp = 1,
  zlim=c(0,3))

# If we don't use a t-test
##########################
# Unclustered
image(
  cbind(all2.1.2[c(dom_sams2.r4, ber_sams2.r4),], sidebar),
  col=redblue100,axes=F,main=glue("chr2:142940000-143220000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

# Clustered
clst_mat <- all2.1.2[c(dom_sams2.r4, ber_sams2.r4),]
clst_mat <- clst_mat[,hmo]
image(
  cbind(clst_mat, sidebar),
  col=redblue100,axes=F,main=glue("chr2:142940000-143220000, Cohort Order, Clustered"), asp = 1,
  zlim=c(0,3))
##########################

# Add RJF

cls2 <- c(cls, rep(0.5,length(rjf_sams2.r4)))
sidebar<-cbind(cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,
               cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,
               cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2,cls2)
sidebar <- sidebar*3

hmo_dom <- heatmap(all2.1.2[c(dom_sams2.r4),r4.roi])$rowInd
hmo_ber <- heatmap(all2.1.2[c(ber_sams2.r4),r4.roi])$rowInd
hmo_ber <- hmo_ber + length(hmo_dom)
hmo_rjf <- heatmap(all2.1.2[c(rjf_sams2.r4),r4.roi])$rowInd
hmo_rjf <- hmo_rjf + length(c(hmo_dom,hmo_ber))
#############
# Unclustered
image(
  cbind(all2.1.2[c(dom_sams2.r4, ber_sams2.r4,rjf_sams2.r4),r4.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr2:142940000-143220000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

# Clustered
clst_mat <- all2.1.2[c(dom_sams2.r4, ber_sams2.r4,rjf_sams2.r4),r4.roi]
clst_mat <- clst_mat[,hmo]
image(
  cbind(clst_mat, sidebar),
  col=redblue100,axes=F,main=glue("chr2:142940000-143220000, Cohort Order, Clustered"), asp = 1,
  zlim=c(0,3))

clst_mat <- all2.1.2[c(hmo_dom, hmo_ber,hmo_rjf),r4.roi]
clst_mat <- clst_mat[,hmo]
image(
  cbind(clst_mat, sidebar),
  col=redblue100,axes=F,main=glue("chr2:142940000-143220000, Cohort Order, Clustered"), asp = 1,
  zlim=c(0,3))

```

## Sweep Region 5 (chr3:52840000-52900000)

### T-test (dom vs ber) (chr3:52840000-52900000)
```{r}
# Perform a differential genotype test between domestic and bermuda
sam_ann_mat3.r5.df <- as.data.frame(sam_ann_mat3.1)
dom_sams3.r5 <- sam_ann_mat3.r5.df %>%
  filter(cohort == 1) %>% row.names()
ber_sams3.r5 <- sam_ann_mat3.r5.df %>%
  filter(cohort == 2) %>% row.names()

# Collect the region specific snps
r.gg5.pos <- all_df3 %>%
  filter(POS %in% colnames(dom3.1.2)) %>%
  select(CHROM,POS,CHROM_gg4,POS_gg4) %>% unique() %>%
  arrange(POS_gg4) %>%
  filter(POS_gg4 >= 52840000) %>% filter(POS_gg4 <= 52900000) %>%
  select(POS) %>% unlist() %>% unname() %>% as.character()

# Collect region specific locations
dom3.r5.2 <- all3.1.2[dom_sams3.r5,r.gg5.pos]
ber3.r5.2 <- all3.1.2[ber_sams3.r5,r.gg5.pos]
all3.r5.2 <- all3.1.2[c(dom_sams3.r5, ber_sams3.r5),r.gg5.pos]
row.names(all3.r5.2)
dim(dom3.r5.2)
dim(ber3.r5.2)
dim(all3.r5.2)

# T-test
t.ber.dom  <- rep(0,dim(all3.r5.2)[2])
cls <- c(rep(0, dim(dom3.r5.2)[1]), rep(1, dim(ber3.r5.2)[1]))
t.ber.dom <- mt.teststat(t(all3.r5.2), cls, test="wilcoxon")
qqnorm(t.ber.dom); qqline(t.ber.dom)
qqnorm(t.ber.dom, ylim = c(-6.5,6.5), xlim = c(-4,4)); qqline(t.ber.dom)
```

### Variants with Positive T-score (chr3:52840000-52900000)
```{r}
# Positive t-scores
t.df.p <- data.frame('P' = colnames(all3.r5.2), 't_val' = t.ber.dom) %>%
  filter(t_val > 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.p2 <- t.df.p %>% filter(!is.na(t_val))
t.df.p2 <- t.df.p2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))

image(all3.r5.2, col = redblue100)
image(all3.r5.2[,t.df.p2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr3') %>%
  filter(POS_gg4 >= 52840000) %>% filter(POS_gg4 <= 52900000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.p2 <- t.df.p2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.p <- t.df.p2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr3') %>% 
  filter(POS %in% t.sig.p) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Variants with Negative T-score (chr3:52840000-52900000)
```{r}
# Negative t-scores
t.df.n <- data.frame('P' = colnames(all3.r5.2), 't_val' = t.ber.dom) %>%
  filter(t_val < 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.n2 <- t.df.n %>% filter(!is.na(t_val))
t.df.n2 <- t.df.n2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all3.r5.2, col = redblue100)
image(all3.r5.2[,t.df.n2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr3') %>%
  filter(POS_gg4 >= 52840000) %>% filter(POS_gg4 <= 52900000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.n2 <- t.df.n2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.n <- t.df.n2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr3') %>% filter(POS %in% t.sig.n) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Visualize the t-scores by location (chr3:52840000-52900000)
```{r}
# Bind dfs
r5.t.df <- rbind(t.df.p2, t.df.n2) %>%
  mutate(t_sign = ifelse(t_val >0, 'pos', 'neg')) %>%
  mutate(t_val_abs = abs(t_val)) %>%
  mutate(del = ifelse(grepl('del', SIFT), 'del', 'tol'))

r5.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(52840000,52900000) +
  theme_bw()

r5.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(52840000,52900000) +
  theme_bw() +
  facet_wrap(~ensembl_geneid + var_type)

# Visualize gg4 position by t scores
r5.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  xlim(52840000,52900000) +
  facet_wrap(~var_type)

# Relavent genes (geneid)
r5.t.df %>%
  #filter(ensembl_geneid != '') %>%
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0.95,1) +
  xlim(52840000,52900000) +
  facet_wrap(~var_type)
```

### Examine the variantions of interest (chr3:52840000-52900000)
```{r}
r5.roi <- r5.t.df %>%
  #filter(ensembl_geneid != '') %>%
  filter(fdr <= 0.05) %>%
  select(P) %>% unique() %>% unlist() %>% unname()

# Ann
sidebar<-cbind(cls,cls,cls,cls)
sidebar <- sidebar*3

hmo <- heatmap(all3.1.2[c(dom_sams3.r5, ber_sams3.r5),r5.roi])$colInd

# Unclustered
image(
  cbind(all3.1.2[c(dom_sams3.r5, ber_sams3.r5),r5.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr3:52840000-52900000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

# Clustered
clst_mat <- all3.1.2[c(dom_sams3.r5, ber_sams3.r5),r5.roi]
clst_mat <- clst_mat[,hmo]
image(
  cbind(clst_mat, sidebar),
  col=redblue100,axes=F,main=glue("chr3:52840000-52900000, Cohort Order, Clustered"), asp = 1,
  zlim=c(0,3))
```

## Sweep Region 6 (chr3:92900000-92960000)

### T-test (dom vs ber) (chr3:92900000-92960000)
```{r}
# Perform a differential genotype test between domestic and bermuda
sam_ann_mat3.r6.df <- as.data.frame(sam_ann_mat3.1)
dom_sams3.r6 <- sam_ann_mat3.r6.df %>%
  filter(cohort == 1) %>% row.names()
ber_sams3.r6 <- sam_ann_mat3.r6.df %>%
  filter(cohort == 2) %>% row.names()

# Collect the region specific snps
r.gg5.pos <- all_df3 %>%
  filter(POS %in% colnames(dom3.1.2)) %>%
  select(CHROM,POS,CHROM_gg4,POS_gg4) %>% unique() %>%
  arrange(POS_gg4) %>%
  filter(POS_gg4 >= 92900000) %>% filter(POS_gg4 <= 92960000) %>%
  select(POS) %>% unlist() %>% unname() %>% as.character()

# Collect region specific locations
dom3.r6.2 <- all3.1.2[dom_sams3.r6,r.gg5.pos]
ber3.r6.2 <- all3.1.2[ber_sams3.r6,r.gg5.pos]
all3.r6.2 <- all3.1.2[c(dom_sams3.r6, ber_sams3.r6),r.gg5.pos]
row.names(all3.r6.2)
dim(dom3.r6.2)
dim(ber3.r6.2)
dim(all3.r6.2)

# T-test
t.ber.dom  <- rep(0,dim(all3.r6.2)[2])
cls <- c(rep(0, dim(dom3.r6.2)[1]), rep(1, dim(ber3.r6.2)[1]))
t.ber.dom <- mt.teststat(t(all3.r6.2), cls, test="wilcoxon")
qqnorm(t.ber.dom); qqline(t.ber.dom)
qqnorm(t.ber.dom, ylim = c(-6.5,6.5), xlim = c(-4,4)); qqline(t.ber.dom)
```

### Variants with Positive T-score (chr3:92900000-92960000)
```{r}
# Positive t-scores
t.df.p <- data.frame('P' = colnames(all3.r6.2), 't_val' = t.ber.dom) %>%
  filter(t_val > 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.p2 <- t.df.p %>% filter(!is.na(t_val))
t.df.p2 <- t.df.p2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))

image(all3.r6.2, col = redblue100)
image(all3.r6.2[,t.df.p2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr3') %>%
  filter(POS_gg4 >= 92900000) %>% filter(POS_gg4 <= 92960000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.p2 <- t.df.p2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.p <- t.df.p2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr3') %>% 
  filter(POS %in% t.sig.p) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Variants with Negative T-score (chr3:92900000-92960000)
```{r}
# Negative t-scores
t.df.n <- data.frame('P' = colnames(all3.r6.2), 't_val' = t.ber.dom) %>%
  filter(t_val < 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.n2 <- t.df.n %>% filter(!is.na(t_val))
t.df.n2 <- t.df.n2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all3.r6.2, col = redblue100)
image(all3.r6.2[,t.df.n2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr3') %>%
  filter(POS_gg4 >= 92900000) %>% filter(POS_gg4 <= 92960000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.n2 <- t.df.n2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.n <- t.df.n2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr3') %>% filter(POS %in% t.sig.n) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Visualize the t-scores by location (chr3:92900000-92960000)
```{r}
# Bind dfs
r6.t.df <- rbind(t.df.p2, t.df.n2) %>%
  mutate(t_sign = ifelse(t_val >0, 'pos', 'neg')) %>%
  mutate(t_val_abs = abs(t_val)) %>%
  mutate(del = ifelse(grepl('del', SIFT), 'del', 'tol'))

r6.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(92900000,92960000) +
  theme_bw()

r6.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(92900000,92960000) +
  theme_bw() +
  facet_wrap(~ensembl_geneid + var_type)

# Visualize gg4 position by t scores
r6.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  xlim(92900000,92960000) +
  facet_wrap(~var_type)

# Relavent genes (geneid)
r6.t.df %>%
  filter(fdr <= 0.1) %>%
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0.90,1) +
  xlim(92900000,92960000) +
  facet_wrap(~var_type)
```

### Examine the variantions of interest (chr3:92900000-92960000)
```{r}
r6.roi <- r6.t.df %>%
  #filter(ensembl_geneid != '') %>%
  filter(fdr <= 0.1) %>%
  select(P) %>% unique() %>% unlist() %>% unname()

# Ann
sidebar<-cbind(cls)
sidebar <- sidebar*3

# Unclustered
image(
  cbind(all3.1.2[c(dom_sams3.r6, ber_sams3.r6),r6.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr3:92900000-92960000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

```

## Sweep Region 7 (chr5:3700000-3960000)

### T-test (dom vs ber) (chr5:3700000-3960000)
```{r}
# Perform a differential genotype test between domestic and bermuda
sam_ann_mat5.r7.df <- as.data.frame(sam_ann_mat5.1)
dom_sams5.r7 <- sam_ann_mat5.r7.df %>%
  filter(cohort == 1) %>% row.names()
ber_sams5.r7 <- sam_ann_mat5.r7.df %>%
  filter(cohort == 2) %>% row.names()

# Collect the region specific snps
r.gg5.pos <- all_df3 %>%
  filter(POS %in% colnames(dom5.1.2)) %>%
  select(CHROM,POS,CHROM_gg4,POS_gg4) %>% unique() %>%
  arrange(POS_gg4) %>%
  filter(POS_gg4 >= 3700000) %>% filter(POS_gg4 <= 3960000) %>%
  select(POS) %>% unlist() %>% unname() %>% as.character()

# Collect region specific locations
dom5.r7.2 <- all5.1.2[dom_sams5.r7,r.gg5.pos]
ber5.r7.2 <- all5.1.2[ber_sams5.r7,r.gg5.pos]
all5.r7.2 <- all5.1.2[c(dom_sams5.r7, ber_sams5.r7),r.gg5.pos]
row.names(all5.r7.2)
dim(dom5.r7.2)
dim(ber5.r7.2)
dim(all5.r7.2)

# T-test
t.ber.dom  <- rep(0,dim(all5.r7.2)[2])
cls <- c(rep(0, dim(dom5.r7.2)[1]), rep(1, dim(ber5.r7.2)[1]))
t.ber.dom <- mt.teststat(t(all5.r7.2), cls, test="wilcoxon")
qqnorm(t.ber.dom); qqline(t.ber.dom)
qqnorm(t.ber.dom, ylim = c(-6.5,6.5), xlim = c(-4,4)); qqline(t.ber.dom)
```

### Variants with Positive T-score (chr5:3700000-3960000)
```{r}
# Positive t-scores
t.df.p <- data.frame('P' = colnames(all5.r7.2), 't_val' = t.ber.dom) %>%
  filter(t_val > 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.p2 <- t.df.p %>% filter(!is.na(t_val))
t.df.p2 <- t.df.p2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))

image(all5.r7.2, col = redblue100)
image(all5.r7.2[,t.df.p2$P], col = redblue100)

# Select the chrom specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr5') %>%
  filter(POS_gg4 >= 3700000) %>% filter(POS_gg4 <= 3960000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.p2 <- t.df.p2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.p <- t.df.p2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr5') %>% 
  filter(POS %in% t.sig.p) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Variants with Negative T-score (chr5:3700000-3960000)
```{r}
# Negative t-scores
t.df.n <- data.frame('P' = colnames(all5.r7.2), 't_val' = t.ber.dom) %>%
  filter(t_val < 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.n2 <- t.df.n %>% filter(!is.na(t_val))
t.df.n2 <- t.df.n2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all5.r7.2, col = redblue100)
image(all5.r7.2[,t.df.n2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr5') %>%
  filter(POS_gg4 >= 3700000) %>% filter(POS_gg4 <= 3960000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.n2 <- t.df.n2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.n <- t.df.n2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr5') %>% filter(POS %in% t.sig.n) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Visualize the t-scores by location (chr5:3700000-3960000)
```{r}
# Bind dfs
r7.t.df <- rbind(t.df.p2, t.df.n2) %>%
  mutate(t_sign = ifelse(t_val >0, 'pos', 'neg')) %>%
  mutate(t_val_abs = abs(t_val)) %>%
  mutate(del = ifelse(grepl('del', SIFT), 'del', 'tol'))

r7.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(3700000,3960000) +
  theme_bw()

r7.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(3700000,3960000) +
  theme_bw() + 
  geom_abline(intercept = 3, slope =0, color = 'blue', size = 0.25) +
  facet_wrap(~ ensembl_geneid + var_type)

# Visualize gg4 position by t scores
r7.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0,1) +
  xlim(3700000,3960000) +
  facet_wrap(~var_type)

r7.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = abs(t_val), color = t_sign)) +
  geom_point() +
  xlim(3700000,3960000)

r7.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = abs(t_val), color = t_sign)) +
  geom_point() +
  xlim(3700000,3960000) +
  facet_wrap(~var_type)
```

### Examine the variantions of interest (chr5:3700000-3960000)
```{r}
r7.roi <- r7.t.df %>%
  filter(ensembl_geneid == 'ENSGALG00000012153') %>%
  filter(var_type %in% c('missense_variant','stop_gained') | grepl('splice', var_type)) %>%
  arrange(desc(POS)) %>%
  filter(VEP_ANN_N == 1) %>%
  #filter(fdr <= 0.1) %>%
  select(P) %>% unique() %>% unlist() %>% unname()

r7.t.df %>%
  filter(ensembl_geneid == 'ENSGALG00000012153') %>%
  filter(var_type %in% c('missense_variant','stop_gained') | grepl('splice', var_type)) %>%
  arrange(desc(POS)) %>%
  filter(VEP_ANN_N == 1) %>%
  select(POS_gg4,t_val,p_val,fdr,REF,ALT,var_type,SIFT)
  
# Ann
sidebar<-cbind(cls)
sidebar <- sidebar*3

# Unclustered
image(
  cbind(all5.1.2[c(dom_sams5.r7, ber_sams5.r7),r7.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr5:3700000-3960000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))

```

## Sweep Region 8 (chr7:19320000-19380000)

### T-test (dom vs ber) (chr7:19320000-19380000)
```{r}
# Perform a differential genotype test between domestic and bermuda
sam_ann_mat7.r8.df <- as.data.frame(sam_ann_mat7.1)
dom_sams7.r8 <- sam_ann_mat7.r8.df %>%
  filter(cohort == 1) %>% row.names()
ber_sams7.r8 <- sam_ann_mat7.r8.df %>%
  filter(cohort == 2) %>% row.names()

# Collect the region specific snps
r.gg5.pos <- all_df3 %>%
  filter(POS %in% colnames(dom7.1.2)) %>%
  select(CHROM,POS,CHROM_gg4,POS_gg4) %>% unique() %>%
  arrange(POS_gg4) %>%
  filter(POS_gg4 >= 19320000) %>% filter(POS_gg4 <= 19380000) %>%
  select(POS) %>% unlist() %>% unname() %>% as.character()

# Collect region specific locations
dom7.r8.2 <- all7.1.2[dom_sams7.r8,r.gg5.pos]
ber7.r8.2 <- all7.1.2[ber_sams7.r8,r.gg5.pos]
all7.r8.2 <- all7.1.2[c(dom_sams7.r8, ber_sams7.r8),r.gg5.pos]
row.names(all7.r8.2)
dim(dom7.r8.2)
dim(ber7.r8.2)
dim(all7.r8.2)

# T-test
t.ber.dom  <- rep(0,dim(all7.r8.2)[2])
cls <- c(rep(0, dim(dom7.r8.2)[1]), rep(1, dim(ber7.r8.2)[1]))
t.ber.dom <- mt.teststat(t(all7.r8.2), cls, test="wilcoxon")
qqnorm(t.ber.dom); qqline(t.ber.dom)
qqnorm(t.ber.dom, ylim = c(-6.5,6.5), xlim = c(-4,4)); qqline(t.ber.dom)
```

### Variants with Positive T-score (chr7:19320000-19380000)
```{r}
# Positive t-scores
t.df.p <- data.frame('P' = colnames(all7.r8.2), 't_val' = t.ber.dom) %>%
  filter(t_val > 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.p2 <- t.df.p %>% filter(!is.na(t_val))
t.df.p2 <- t.df.p2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))

image(all7.r8.2, col = redblue100)
image(all7.r8.2[,t.df.p2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr7') %>%
  filter(POS_gg4 >= 19320000) %>% filter(POS_gg4 <= 19380000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.p2 <- t.df.p2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.p <- t.df.p2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr7') %>% 
  filter(POS %in% t.sig.p) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Variants with Negative T-score (chr7:19320000-19380000)
```{r}
# Negative t-scores
t.df.n <- data.frame('P' = colnames(all7.r8.2), 't_val' = t.ber.dom) %>%
  filter(t_val < 0) %>%
  rowwise() %>%
  mutate(p_val = 2*(1-stats::pnorm(abs(t_val)))) %>%
  ungroup() %>%
  arrange(p_val)

t.df.n2 <- t.df.n %>% filter(!is.na(t_val))
t.df.n2 <- t.df.n2 %>% mutate(fdr = p.adjust(p_val, method = "fdr", n = length(t.ber.dom)))
  
image(all7.r8.2, col = redblue100)
image(all7.r8.2[,t.df.n2$P], col = redblue100)

# Select the chrom 2 specific vars
ch2_var_df <- all_df3 %>%
  filter(CHROM == 'chr7') %>%
  filter(POS_gg4 >= 19320000) %>% filter(POS_gg4 <= 19380000) %>%
  select(POS, POS_gg4, REF, ALT, VEP_ANN_N, var_type, var_impact, gene_symbol, ensembl_geneid, SIFT, protein_domains, exon1, exon2, transcript_type, transcript, ensembl_transcript) %>%
  unique() %>%
  mutate(P = as.character(POS))

# Examine the ranked list of potential variants
t.df.n2 <- t.df.n2 %>%
  left_join(y = ch2_var_df, by = c('P'))

# Select significantly different GT averages based on hypothesis test
t.sig.n <- t.df.n2 %>%
  filter(fdr<=0.05) %>%
  select(P) %>% unlist() %>% unname()
# Visualize those averages
all_df3 %>%
  filter(CHROM == 'chr7') %>% filter(POS %in% t.sig.n) %>%
  select(gene_symbol, var_type) %>%
  table()
```

### Visualize the t-scores by location (chr7:19320000-19380000)
```{r}
# Bind dfs
r8.t.df <- rbind(t.df.p2, t.df.n2) %>%
  mutate(t_sign = ifelse(t_val >0, 'pos', 'neg')) %>%
  mutate(t_val_abs = abs(t_val)) %>%
  mutate(del = ifelse(grepl('del', SIFT), 'del', 'tol'))

r8.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(19320000,19380000) +
  theme_bw()

r8.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = t_val_abs, color = t_sign)) +
  geom_point() + 
  ylim(0,6.5) +
  xlim(19320000,19380000) +
  theme_bw() +
  facet_wrap(~ ensembl_geneid + var_type)

# Visualize gg4 position by t scores
r8.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = abs(t_val), color = t_sign)) +
  geom_point() +
  xlim(19320000,19380000) +
  facet_wrap(~var_type + gene_symbol)

# Visualize gg4 position by t scores
r8.t.df %>%
  #select(POS_gg4, t_val, p_val, fdr, var_type, var_impact, gene_symbol, ensembl_geneid)
  ggplot(aes(x = POS_gg4, y = 1-fdr, color = t_sign)) +
  geom_point() +
  ylim(0,1) +
  xlim(19320000,19380000) +
  facet_wrap(~var_type + gene_symbol)
```

### Examine the variantions of interest (chr7:19320000-19380000)
```{r}
r8.roi <- r8.t.df %>%
  #filter(ensembl_geneid != '') %>%
  filter(t_val_abs >= 2.5) %>%
  select(P) %>% unique() %>% unlist() %>% unname()

# Ann
sidebar<-cbind(cls)
sidebar <- sidebar*3

hmo <- heatmap(all7.1.2[c(dom_sams7.r8, ber_sams7.r8),r8.roi])$colInd

# Unclustered
image(
  cbind(all7.1.2[c(dom_sams7.r8, ber_sams7.r8),r8.roi], sidebar),
  col=redblue100,axes=F,main=glue("chr7:19320000-19380000, Cohort Order, Unclustered"), asp = 1,
  zlim=c(0,3))
```

# Examine the significantly different mutations by adjacent sequences (Investigation of poly-adenylated tracks)
Conclusion: The polyA hypothesis is not supported
```{r}
library("BSgenome")
library("BSgenome.Ggallus.UCSC.galGal4")


r1.t.df %>%
  select(P, POS, POS_gg4)

r1.t.df$CHROM <- 'chr1'
r1.t.df$SWEEP <- '1'
r1.t.df$SEQ_ps <- getSeq(Ggallus, names = r1.t.df$CHROM, start=r1.t.df$POS_gg4-10, end=r1.t.df$POS_gg4+10,
       strand="+", as.character=TRUE)
r2.t.df$CHROM <- 'chr2'
r2.t.df$SWEEP <- '2'
r2.t.df$SEQ_ps <- getSeq(Ggallus, names = r2.t.df$CHROM, start=r2.t.df$POS_gg4-10, end=r2.t.df$POS_gg4+10,
       strand="+", as.character=TRUE)
r3.t.df$CHROM <- 'chr2'
r3.t.df$SWEEP <- '3'
r3.t.df$SEQ_ps <- getSeq(Ggallus, names = r3.t.df$CHROM, start=r3.t.df$POS_gg4-10, end=r3.t.df$POS_gg4+10,
       strand="+", as.character=TRUE)
r4.t.df$CHROM <- 'chr2'
r4.t.df$SWEEP <- '4'
r4.t.df$SEQ_ps <- getSeq(Ggallus, names = r4.t.df$CHROM, start=r4.t.df$POS_gg4-10, end=r4.t.df$POS_gg4+10,
       strand="+", as.character=TRUE)
r5.t.df$CHROM <- 'chr3'
r5.t.df$SWEEP <- '5'
r5.t.df$SEQ_ps <- getSeq(Ggallus, names = r5.t.df$CHROM, start=r5.t.df$POS_gg4-10, end=r5.t.df$POS_gg4+10,
       strand="+", as.character=TRUE)
r6.t.df$CHROM <- 'chr3'
r6.t.df$SWEEP <- '6'
r6.t.df$SEQ_ps <- getSeq(Ggallus, names = r6.t.df$CHROM, start=r6.t.df$POS_gg4-10, end=r6.t.df$POS_gg4+10,
       strand="+", as.character=TRUE)
r7.t.df$CHROM <- 'chr5'
r7.t.df$SWEEP <- '7'
r7.t.df$SEQ_ps <- getSeq(Ggallus, names = r7.t.df$CHROM, start=r7.t.df$POS_gg4-10, end=r7.t.df$POS_gg4+10,
       strand="+", as.character=TRUE)
r8.t.df$CHROM <- 'chr7'
r8.t.df$SWEEP <- '8'
r8.t.df$SEQ_ps <- getSeq(Ggallus, names = r8.t.df$CHROM, start=r8.t.df$POS_gg4-10, end=r8.t.df$POS_gg4+10,
       strand="+", as.character=TRUE)

allr.t.df <- rbind(r1.t.df,r2.t.df,r3.t.df,r4.t.df,
                   r5.t.df,r6.t.df,r7.t.df,r8.t.df)

# Generate a column for the percentage of A's and T's
allr.t.df$SEQ_ps <- as.character(allr.t.df$SEQ_ps)
a_count <- str_count(allr.t.df$SEQ_ps, pattern = 'A')
tot <- mean(nchar(allr.t.df$SEQ_ps))
allr.t.df$PERCA <- a_count/tot
# By T's
t_count <- str_count(allr.t.df$SEQ_ps, pattern = 'T')
allr.t.df$PERCT <- t_count/tot
# Generate a column for the percentage of G's and C's
g_count <- str_count(allr.t.df$SEQ_ps, pattern = 'G')
allr.t.df$PERCG <- g_count/tot
# By C's
c_count <- str_count(allr.t.df$SEQ_ps, pattern = 'C')
allr.t.df$PERCC <- c_count/tot




hist(allr.t.df$PERCA)
hist(allr.t.df$PERCC)

allr.t.df %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(y = var_type, x = PERCA)) +
  geom_boxplot() +
  facet_wrap(~ SWEEP)

allr.t.df %>%
  filter(t_val_abs >= 3) %>%
  ggplot(aes(x = PERCA, color = SWEEP)) +
  geom_density()

allr.t.df %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(y = var_type, x = PERCC)) +
  geom_boxplot() +
  facet_wrap(~ SWEEP)

allr.t.df %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(x = PERCC, color = SWEEP)) +
  geom_density()

allr.t.df %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(x = PERCA, color = SWEEP)) +
  geom_density()

df_sig <- allr.t.df %>%
  filter(fdr <= 0.05)

hist(df_sig$PERCA)
hist(df_sig$PERCC)

df_sig %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(y = var_type, x = PERCA)) +
  geom_boxplot() +
  facet_wrap(~ SWEEP)

df_sig %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(x = PERCA, color = SWEEP)) +
  geom_density()

df_sig %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(y = var_type, x = PERCC)) +
  geom_boxplot() +
  facet_wrap(~ SWEEP)

df_sig %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(x = PERCC, color = SWEEP)) +
  geom_density()

df_sig %>%
  filter(fdr <= 0.05) %>%
  ggplot(aes(x = PERCA, color = SWEEP)) +
  geom_density()

df_sig %>%
  filter(PERCA > 0.6)


```

# Session Info
```{r Sesh}
warnings()
session_info()
```

`r knitr::knit_exit()`